todo:
- check coding
- check RT tablet computer
- item info for 3Ds
- deal with repeated measures in IRT
X ability by age 
X write out scores

```{r setup}
source(here::here("assessment/colombia/stage2/_setup.R"))

trials_coded <- read_rds(here("assessment/colombia/stage2/data_processed/trials_coded.rds"))
participants <- read_rds(here("assessment/colombia/stage2/data_processed/participants.rds"))
```

Munge data.

```{r}
mr <- trials_coded |>
  filter(task_id == "mental-rotation") |>
  select(user_id, run_id, item, response, is_correct, rt, age, server_timestamp) |>
  mutate(rt = as.numeric(rt),
         stim_type = if_else(str_count(item) == 5, "2d", "3d"),
         stim_set = str_sub(item, 1, 1) |> fct_recode("rabbit" = "r", "duck" = "d"),
         stim_set = if_else(stim_type == "2d", glue("2d ({stim_set})"), "3d"),
         angle = item |> str_sub(-3) |> as.numeric(),
         angle = if_else(angle > 180, 360 - angle, angle),
         .after = item) |>
  mutate(rt = rt / 1000)
```

```{r}
mr |> distinct(stim_type, stim_set, angle, item) |> count(stim_type, stim_set, angle)
```

# N trials

```{r}
mr_summary <- mr |> 
  group_by(user_id, age) |>
  summarise(prop_correct =  mean(is_correct), 
            n = n()) |>
  ungroup()
```

```{r}
ggplot(mr_summary, aes(x = n)) + 
  geom_histogram(bins = 10)
```

Classic confounding of sum score and n trials.

```{r}
ggplot(mr_summary, aes(x = n, y = prop_correct)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```



# Accuracy

```{r}
mr_subject_type <- mr |>
  group_by(user_id, stim_type, angle) |>
  summarise(items = n_distinct(item),
            prop_correct = mean(is_correct, na.rm = TRUE))

mr_type <- mr_subject_type |>
  group_by(stim_type, angle) |>
  summarise(accuracy = mean(prop_correct),
            sd = sd(prop_correct),
            n = n(),
            se = sd / sqrt(n))
```

```{r}
ggplot(mr_subject_type, aes(x = angle, y = prop_correct, colour = stim_type)) +
  facet_grid(cols = vars(stim_type), scales = "free_x", space = "free_x") +
  geom_jitter(alpha = 0.5, size = 0.8) +
  geom_pointrange(aes(y = accuracy, ymin = accuracy - 1.96 * se, ymax = accuracy + 1.96 * se),
                  data = mr_type) +
  geom_line(aes(y = accuracy), data = mr_type) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_continuous(breaks = unique(mr_subject_type$angle)) +
  scale_colour_ptol(guide = "none") +
  labs(x = "Angle", y = "Proportion correct responses", colour = "")

#ggsave("plots/mr_accuracy_type.png", width = 6, height = 3.5)
```


```{r}
mr_subject_set <- mr |>
  group_by(user_id, stim_type, stim_set, angle) |>
  summarise(items = n_distinct(item),
            prop_correct = mean(is_correct, na.rm = TRUE))

mr_set <- mr_subject_set |>
  group_by(stim_type, stim_set, angle) |>
  summarise(accuracy = mean(prop_correct),
            sd = sd(prop_correct),
            n = n(),
            se = sd / sqrt(n))
```

```{r}
ggplot(mr_subject_set, aes(x = angle, y = prop_correct, colour = stim_type)) +
  facet_grid(cols = vars(stim_set), scales = "free_x", space = "free_x") +
  geom_jitter(alpha = 0.5, size = 0.8) +
  geom_pointrange(aes(y = accuracy, ymin = accuracy - 1.96 * se, ymax = accuracy + 1.96 * se),
                  data = mr_set, position = position_dodge(width = 10)) +
  geom_line(aes(y = accuracy), data = mr_set) +
  scale_x_continuous(breaks = unique(mr_subject_set$angle)) +
  scale_colour_ptol(guide = "none") +
  labs(x = "Angle", y = "Proportion correct responses", colour = "")
ggsave("plots/mr_accuracy_set.png", width = 11, height = 3.5)
```

# RTs

```{r}
mr_subject_type_rt <- mr |>
  group_by(user_id, stim_type, angle, is_correct) |>
  summarise(items = n_distinct(item),
            median_rt = median(rt, na.rm = TRUE)) |>
  mutate(correct = if_else(is_correct, "correct", "incorrect"))

mr_type_rt <- mr_subject_type_rt |>
  group_by(stim_type, angle, correct) |>
  summarise(rt = mean(median_rt),
            sd = sd(median_rt),
            n = n(),
            se = sd / sqrt(n))
```

```{r}
ggplot(mr_subject_type_rt, aes(x = angle, y = median_rt, colour = stim_type)) +
  facet_grid(cols = vars(stim_type), rows = vars(correct), scales = "free_x", space = "free_x") +
  geom_jitter(alpha = 0.5, size = 0.8) +
  geom_pointrange(aes(y = rt, ymin = rt - 1.96 * se, ymax = rt + 1.96 * se),
                  data = mr_type_rt) +
  geom_line(aes(y = rt), data = mr_type_rt) +
  scale_x_continuous(breaks = unique(mr_subject_type$angle)) +
  scale_colour_ptol(guide = "none") +
  scale_y_log10() + 
  labs(x = "Angle", y = "Response time (seconds)", colour = "")
ggsave("plots/mr_rt_type.png", width = 6, height = 5)
```

```{r}
mr_subject_set_rt <- mr |>
  group_by(user_id, stim_type, stim_set, angle, is_correct) |>
  summarise(items = n_distinct(item),
            median_rt = median(rt, na.rm = TRUE)) |>
  mutate(correct = if_else(is_correct, "correct", "incorrect"))

mr_set_rt <- mr_subject_set_rt |>
  group_by(stim_type, stim_set, angle, correct) |>
  summarise(rt = mean(median_rt),
            sd = sd(median_rt),
            n = n(),
            se = sd / sqrt(n))
```

```{r}
ggplot(mr_subject_set_rt, aes(x = angle, y = median_rt, colour = stim_type)) +
  facet_grid(cols = vars(stim_set), row = vars(correct), scales = "free_x", space = "free_x") +
  geom_jitter(alpha = 0.5, size = 0.8) +
  geom_pointrange(aes(y = rt, ymin = rt - 1.96 * se, ymax = rt + 1.96 * se),
                  data = mr_set_rt) +
  geom_line(aes(y = rt), data = mr_set_rt) +
  scale_x_continuous(breaks = unique(mr_subject_set_rt$angle)) +
  scale_colour_ptol(guide = "none") +
  scale_y_log10() + 
  labs(x = "Angle", y = "Response time (seconds)", colour = "")
ggsave("plots/mr_rt_set.png", width = 10, height = 5)
```

# IRT models

Prepare data. 

Repeated measures have constrained parameters (both difficulty and discrimination). 

```{r}
mr_first_runs <- mr |>
  group_by(user_id) |>
  filter(server_timestamp == min(server_timestamp)) |>
  select(user_id, run_id)

mr_wide <- mr |>
  inner_join(mr_first_runs) |>
  mutate(is_correct = as.numeric(is_correct)) |>
  select(user_id, item, is_correct) |> 
  group_by(user_id, item) |>
  mutate(instance = seq_along(item),
         item_inst = glue("{item}_{instance}")) |> 
  ungroup() |> 
  select(-item, -instance) |> 
  pivot_wider(names_from = "item_inst", values_from = "is_correct")

mr_mat <- as.matrix(select(mr_wide, -user_id))
rownames(mr_mat) <- mr_wide$user_id

# remove columns with no variance
mr_cols <- colMeans(mr_mat, na.rm=TRUE) 

mr_mat <- mr_mat[,mr_cols > 0 & mr_cols < 1]
```

Fit models. 

```{r psycho-models_1pl, echo=F, eval = TRUE}
set.seed(1234)

items <- mr$item |> unique()

constraints <- lapply(items, \(i) {
  matched <- which(str_detect(colnames(mr_mat), glue("^{i}_")))
  if (length(matched) > 1) {
    glue("({paste(matched, collapse=',')},a1), ({paste(matched, collapse=',')},d)")
  }
}) |> keep( ~ !is.null(.)) |> paste(collapse = ", ")

mod_str <- glue("F = 1-{ncol(mr_mat)}
                 CONSTRAIN = {constraints}")

fit_mirt_mr <- function(data_mat, item_type) {
  mod <- mirt(data_mat, mod_str, itemtype = item_type, verbose = TRUE, guess = .5,
              technical = list(NCYCLES = 1000))
  
  mod_coefs <- coef(mod, simplify = TRUE)$items |> as_tibble(rownames = "item")
  
  scores <- fscores(mod, method = "MAP", verbose = FALSE)
  mod_fscores <- tibble(user_id = rownames(data_mat), ability = scores[,1])
  
  list(model = mod, coefs = mod_coefs, fscores = mod_fscores)
}

mr_1pl <- fit_mirt_mr(mr_mat, "Rasch")
write_rds(mr_1pl, here("assessment/colombia/stage2/task_analyses/mental_rotation/fits/mr_1pl.rds"))

mr_2pl <- fit_mirt_mr(mr_mat, "2PL")
write_rds(mr_2pl, here("assessment/colombia/stage2/task_analyses/mental_rotation/fits/mr_2pl.rds"))
```

Rasch model wins.

```{r}
anova(mr_1pl$model, mr_2pl$model)
mr_best <- mr_1pl
```

Coefficients. 

```{r}
coefs_1pl_info <- mr_best$coefs |>
  separate_wider_delim(item, names = c("item", "instance"), delim = "_") |> 
  select(-instance) |> 
  distinct() |> 
  left_join(select(mr, item, angle, stim_type, stim_set) |> distinct()) 

ggplot(coefs_1pl_info, 
       aes(x = angle, y = d, col = stim_set)) + 
  geom_jitter(alpha = .5, width = .5) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~stim_type, scales = "free_y")
```

```{r}
mr_summary <- mr_summary |> left_join(mr_best$fscores)
GGally::ggpairs(mr_summary |> select(-user_id))
```

```{r}
ggplot(mr_summary, aes(x = prop_correct, y = ability)) + 
  geom_point(aes(size = n), alpha = .5) + 
  geom_smooth(method = "lm")
```

# Write out parameters

```{r}
mr_scores <- mr_best$fscores |>
  rename(score = ability) |>
  mutate(task = "mental-rotation", score_type = "ability (1PL)")

write_rds(mr_scores, here("assessment/colombia/stage2/scores/mr_scores.rds"))
```

## Add item parameters to item-bank
```{r}
mr_1pl <- read_rds(here("assessment/colombia/stage2/task_analyses/mental_rotation/fits/mr_1pl.rds")) 
mr_item_bank <- read_csv(here("item-banks/mental-rotation-item-bank.csv")) |>
  mutate(item_id = item |> str_to_lower() |> str_replace('-', ''))

itpars <- mr_1pl$coefs |> 
  mutate(item_id = item |> str_remove("_1$") |> str_remove("_2$"),
         difficulty = -d) |>
  select(-item, -d) |> distinct() 
  

mr_item_bank <- mr_item_bank |> left_join(itpars) |> rename(a = a1)
write_csv(mr_item_bank, here("item-banks/mental-rotation-item-bank_item-params.csv"))
```

