---
title: "LEVANTE Egma"
author: "Fionnuala O'Reilly"
toc: true
number-sections: true
highlight-style: pygments
format:
  html: 
    code-fold: true
---
Required packages

```{r}
#| echo: false

library("tidyverse")
library("glue")
library("here")
library("viridis")
library("purrr")
library("readr")
library("dplyr")
library("tidyr")
library("quarto")
library("forcats")
library("ggplot2")
```

Helper functions

```{r}
#| echo: false

source(here("02_score_data","02_fit_irt","irt_helpers.R"))
source(here::here("plot_settings.R"))
source(here("03_explore_tasks","explore_helper.R"))
```

Get egma data

```{r}
task_data_nested <- read_rds(here(glue("01_fetched_data/task_data_nested.rds")))
glimpse(task_data_nested)

egma <- load_task_data("math")
egma
```

Exploring item group

```{r}
base::table(egma$item_group)
```
Sum scores

```{r}
# 94 items
colnames(egma)

egma_runs <- egma |>
  group_by(site, user_id, run_id) |>
  summarise(
    correct = mean(correct, na.rm = TRUE),  
    age = mean(age, na.rm = TRUE),          
    n_items = n_distinct(item_uid) # how many unique items were completed
  )

egma_sum_scores <- ggplot(egma_runs, aes(x = age, y = correct)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "re")) +
  ylim(0, 1) +
  facet_wrap(~site) +
  labs(
    title = "Age-related accuracy, smoothed trends by site",
    x = "Age (years)",
    y = "Proportion correct",
    caption = "Note: Each point represents a participant run; smoothed trend by site"
  )

egma_sum_scores

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "egma_sum_scores.png"),
  plot = egma_sum_scores,
  width = 10, height = 6, dpi = 300
)
```

Accuracy on item groups

```{r}
egma_cat_runs <- egma |>
  group_by(site, user_id, run_id, item_group) |>
  summarise(
    correct = mean(correct, na.rm = TRUE),
    age = mean(age, na.rm = TRUE),
    n_items = n_distinct(item_uid),
    .groups = "drop"
  )

egma_counts <- egma_cat_runs |>
  group_by(site, item_group) |>
  summarise(n = n(), .groups = "drop")
```

Plot

```{r fig.height =8, fig.width = 10}

ggplot(egma_cat_runs, aes(x = age, y = correct)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  geom_text(data = egma_counts,
          aes(x = Inf, y = Inf, label = paste0("n = ", n)),
          hjust = -0.1, vjust = 1,
          size = 3, 
          inherit.aes = FALSE) +
  ylim(0, 1) +
  facet_grid(site ~ item_group) +
  labs(title = "EGMA accuracy by age, site, and category",
       x = "Age (years)",
       y = "Proportion correct")

```

Cleaner graph 

```{r}
egma_type <- egma |>
  group_by(site, user_id, run_id, item_group) |>
  summarise(correct = mean(correct, na.rm = TRUE),
            age = mean(age, na.rm = TRUE))

ggplot(egma_type, aes(x = age, y = correct, col = item_group)) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~site)
```

Difficulty

```{r}
# Proportion correct per item per site
item_difficulty <- egma |>
  group_by(site, item_uid) |>
  summarise(p_correct = mean(correct, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = site, values_from = p_correct)

# Man accuracy across sites
egma_difficulty_raw <- item_difficulty |>
  mutate(mean_accuracy = rowMeans(across(-item_uid), na.rm = TRUE))

# Reorder items from easiest to hardest
egma_difficulty_raw <- egma_difficulty_raw |>
  mutate(item_uid = fct_reorder(item_uid, mean_accuracy))
```

Plot

```{r fig.height=20, fig.width=10}
ggplot(egma_difficulty_raw, aes(x = item_uid, y = mean_accuracy, fill = mean_accuracy)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "EGMA item difficulty (easiest to hardest)",
    x = "Item (UID)",
    y = "Mean proportion correct",
    fill = "Accuracy"
  ) +
  ylim(0, 1) 

# Note: The 'easiest' items are based on the proportion of correct responses, which can be misleading if only a few participants attempted them. For example, two fraction items appear among the easiest, even though fractions are generally the most difficult category—this is because the few participants who saw those items answered them correctly. IRT model accounts for this as it estimates item difficulty and discrimination while adjusting for participant ability.
```

Difficulty by item_group

```{r}
egma_groups <- egma |>
  select(item_uid, item_group) |>
  distinct()

egma_difficulty_grouped <- egma_difficulty_raw |>
  mutate(item_uid = as.character(item_uid)) |>
  left_join(egma_groups, by = "item_uid")

# Mean accuracy by item group
egma_group_summary <- egma_difficulty_grouped |>
  group_by(item_group) |>
  summarise(
    mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n_items = n(),
    .groups = "drop"
  ) |>
  arrange(desc(mean_accuracy))

# Bar graph
ggplot(egma_group_summary, aes(x = fct_reorder(item_group, mean_accuracy), y = mean_accuracy, fill = mean_accuracy)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Average difficulty by item group (EGMA)",
    x = "Item group",
    y = "Mean proportion correct",
    fill = "Accuracy"
  ) +
  ylim(0, 1)

```

Box plot - spread of accuracy across categories

```{r}
ggplot(egma_difficulty_grouped, 
       aes(x = fct_reorder(item_group, mean_accuracy), y = mean_accuracy, fill = item_group)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Distribution of item difficulty by group (EGMA)",
    x = "Item group", y = "Mean accuracy"
  ) 

# Dots
egma_accuracy_raw <- 
  ggplot(egma_difficulty_grouped, 
         aes(x = fct_reorder(item_group, mean_accuracy), y = mean_accuracy, 
             colour = item_group)) +
  geom_jitter() +
  coord_flip() +
  labs(
    title = "Distribution of item difficulty by group (EGMA)",
    x = "Item group", y = "Mean accuracy"
  ) 

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "egma_accuracy_raw.png"),
  plot = egma_accuracy_raw,
  width = 10,
  height = 6,
  dpi = 300
)
```

Reaction times

```{r}
summary(egma$rt_numeric)

egma_trimmed <- egma |> 
  filter(rt_numeric > 200, rt_numeric < 10000)

ggplot(egma_trimmed, aes(x = rt_numeric, fill = correct)) +
  facet_wrap(~site, scales = "free_y") +
  geom_histogram(binwidth = 100) +
  labs(
    title = "Histogram of reaction times by site",
    x = "Reaction time (ms)",
    y = "Count"
  )

# Check outliers
summary(egma$rt_numeric)
summary(egma_trimmed$rt_numeric)
```



# IRT 
Load IRT results
ToDo: fit bifactor model with each subtask

```{r}
# ToDo: update to use model_registry/math/ dir?
best_multigroup <- readRDS(here("02_scoring_outputs", "old", "irt_outputs", "multigroup_best_models.rds"))
multigroup_scores <- readRDS(here("02_scoring_outputs", "scores", "old", "multigroup_scores.rds"))

best_multigroup_coefs <- best_multigroup |>
  select(item_task, item_type, invariance_type, coefs) |>
  unnest(coefs)

best_coefs <- readRDS(here("02_scoring_outputs", "best_coefs.rds")) |> 
  filter(task=="math")
#all_coefs <- readRDS(here("02_scoring_outputs", "all_coefs.rds"))

```

Multigroup models

Developmental trends in egma math ability estimates by site

This plot shows how IRT-derived ability estimates (theta scores, i.e. estimated latent ability) from the Egma task vary with age across different sites.
```{r}

run_ages <- egma |>
  select(site, run_id, user_id, age) |>
  distinct()

multigroup_scores_egma <- multigroup_scores |>
  mutate(site = ifelse(site=="ca_pilot", "pilot_western_ca",
                       ifelse(site=="co_pilot", "pilot_uniandes_co", 
                              ifelse(site=="de_pilot", "pilot_mpieva_de", site)))) |>
  filter(item_task == "math", site != "us_pilot") |>
  select(site, item_task, user_id, run_id, metric_type, metric_value) |>
  left_join(run_ages)

glimpse(multigroup_scores_egma)

egma_irt <- ggplot(multigroup_scores_egma, aes(x = age, y = metric_value, col = site)) + 
  geom_point() + geom_smooth() +
  labs(
    title = "Developmental trends in EGMA ability by site",
    subtitle = "IRT-based theta estimates smoothed with GAM",
    x = "Age (years)",
    y = "Ability (IRT Theta Score)",
    colour = "Site"
  )

egma_irt
ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "egma_irt.png"),
  plot = egma_irt,
  width = 6, height = 4, dpi = 300
)

```

Extract the coefficients  and filter for difficulty

```{r}
# Extract coefficients
egma_difficulty <- best_multigroup_coefs |> 
  filter(item_task == "math") |> 
  select(site, item, d)

# Compute average difficulty
egma_mean_difficulty <- egma_difficulty |> 
  group_by(item) |> 
  summarise(mean_d = mean(d, na.rm = TRUE), .groups = "drop")

# Join with item group
egma_groups <- egma |> 
  select(item_uid, item_group) |> 
  distinct() |> 
  mutate(item = glue("{item_uid}_1"))  

egma_difficulty_grouped <- left_join(egma_mean_difficulty, egma_groups, by = "item")

egma_difficulty_grouped |> filter(is.na(item_group))
```

Item difficulty by group

```{r}
ggplot(egma_difficulty_grouped, aes(x = fct_reorder(item_group, -mean_d), y = mean_d, fill = item_group)) +
  geom_boxplot(alpha = 0.8) +
  coord_flip() +
  labs(
    title = "Distribution of EGMA Item Difficulty by Item Group",
    x = "Item Group",
    y = "Mean IRT Difficulty"
  ) 

# Violin plot
ggplot(egma_difficulty_grouped, aes(x = fct_reorder(item_group, -mean_d), y = mean_d, fill = item_group)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.1, fill = "white", outlier.size = 0.5) +
  coord_flip() +
  labs(
    title = "Distribution of EGMA item difficulty by group",
    x = "Item group",
    y = "Mean IRT difficulty"
  )

# Dots
ggplot(egma_difficulty_grouped, aes(x = fct_reorder(item_group, -mean_d), y = mean_d, colour = item_group)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.7) +
  coord_flip() +
  labs(
    title = "Item difficulty by group (individual points)",
    x = "Item group",
    y = "Mean IRT difficulty"
  )

```


# Number-to-line vs. line-to-number

```{r, eval=T}
# FixMe - egma_difficulty_grouped no longer includes site group, and site names are outdated
line_trials <- egma_difficulty_grouped |>
  filter(str_detect(item_uid, "line") | str_detect(item_uid, "slider")) |>
  separate(item_uid, into = c("item", "trial_type", "point","scale"), sep = "_") |>
  select(-mean_accuracy) |>
  pivot_longer(ca_pilot:de_pilot, names_to = "site", values_to = "correct") |>
  group_by(point, scale, site) |>
  mutate(match = n() == 2) |>
  filter(match) |>
  select(-item, -item_group, -match) |>
  ungroup() |>
  pivot_wider(names_from = "trial_type", values_from = "correct", id_cols = c("site", "point","scale"))


ggplot(line_trials, aes(x = line, y = slider)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~site)
```

Item parameters

```{r}
params <- best_multigroup_coefs |>
  filter(item_task == "math", str_detect(item, "line|slider")) |>
  separate(item, into = c("item", "trial_type", "point","scale","num"), sep = "_") |>
  filter(site == "de_pilot") |>
  select(item, trial_type, point, scale, d) |>
  pivot_wider(names_from = "trial_type", values_from = "d")

ggplot(params, aes(x = -line, y = -slider, col = scale)) + 
  geom_point()+ 
  # geom_smooth(method = "lm", aes(group = 1)) +
  geom_smooth(method = "lm", aes(group = 1),
              data = filter(params, !(scale=="100" & point == "9"))) +
  ggrepel::geom_text_repel(aes(label = point), size = 3) + 
  xlab("Line-to-number 4AFC difficulty") + 
  ylab("Number-to-line slider difficulty")

with(filter(params, !(scale=="100" & point == "9")), 
     cor.test(line, slider))
with(params,  cor.test(line, slider))
```

```{r, eval=F}
filter(egma, item_uid == "math_line_9_100") |> View()
```

# Examine Dimensionality

Note: 34% of responses are slower than 5s, yet have high accuracy. 12% are slower than 10s; I'm going to filter only trials >20s (2.6% of trials), and those <200ms.
We fit a 2PL model to these data, as well as exploratory 2- and 3-factor models, a CFA with each subtask fitted separately, and a bifactor model with each subtask, and a general factor. Model comparisons summarized below.

```{r, eval=F}
# exploratory 2-4 factor models, and bifactor model with each subtask
# nrow(egma |> filter(rt_numeric > 5000)) / nrow(egma) # 34% of responses are slower than 5s, & high accuracy

egma_sorted <- egma |> filter(rt_numeric > 200, rt_numeric < 20000) |>
  arrange(site, user_id, run_id, item_group, item_uid)

egma_wide <- egma_sorted |> select(site, user_id, age, run_id, item_uid, correct) |>
  pivot_wider(id_cols = c(site, user_id, run_id, age), names_from = item_uid, values_from = correct)

egma_wide_rt <- egma_sorted |> select(site, user_id, age, run_id, item_uid, rt_numeric) |>
  pivot_wider(id_cols = c(site, user_id, run_id, age), names_from = item_uid, values_from = rt_numeric)

egma_demo <- egma_wide |> select(site, user_id, age, run_id)
egma_dat <- egma_wide |> select(-site, -user_id, -age, -run_id)# |> data.matrix()
egma_dat_rt <- egma_wide_rt |> select(-site, -user_id, -age, -run_id)

one_value_cols <- names(egma_dat)[sapply(egma_dat, function(x) {
  length(unique(na.omit(x))) == 1
})]

egma_dat <- data.matrix(egma_dat |> select(-all_of(one_value_cols)))
rownames(egma_dat) = egma_demo$run_id

egma_dat_rt <- data.matrix(egma_dat_rt |> select(-all_of(one_value_cols)))
rownames(egma_dat_rt) = egma_demo$run_id

require("diffIRT")
mod_drt <- diffIRT(egma_dat_rt, egma_dat)
#mirt(egma_dat, 1, itemtype='2PL', verbose=TRUE,  technical=list(NCYCLES=3000))

mod_2pl <- mirt(egma_dat, 1, itemtype='2PL', verbose=TRUE,  technical=list(NCYCLES=3000))
mod_2pl_2f <- mirt(egma_dat, 2, itemtype='2PL', verbose=TRUE,  technical=list(NCYCLES=3000))
mod_2pl_3f <- mirt(egma_dat, 3, itemtype='2PL', verbose=TRUE,  technical=list(NCYCLES=3000))

egma_dat_sorted <- egma_dat[,order(colnames(egma_dat))]

cfa <- "add = 1-49
comp = 50-69
frac = 70-83
iden = 84-123
line = 124-194
miss = 195-214
mult = 215-256
slid = 257-295
sub = 296-345"
mod_cfa <- mirt(egma_dat_sorted, cfa, itemtype = "2PL", technical = list(NCYCLES=2000))

bifact <- "gen = 1-345
add = 1-49
comp = 50-69
frac = 70-83
iden = 84-123
line = 124-194
miss = 195-214
mult = 215-256
slid = 257-295
sub = 296-345"
mod_bif <- mirt(egma_dat_sorted, bifact, itemtype = "2PL", technical = list(NCYCLES=2000, MAXQUAD=60000))

save(mod_2pl, mod_2pl_2f, mod_2pl_3f, mod_cfa, mod_bif, 
     file=here::here("03_explore_tasks","math_models.Rdata"))
```

```{r}
load(here::here("03_explore_tasks","math_models.Rdata"))


bind_rows(anova(mod_2pl, mod_2pl_2f),  # AIC prefers 2d, BIC prefers 1d
          anova(mod_2pl_2f, mod_2pl_3f)[2,]) # AIC prefers 3d, BIC prefers 2d



bind_rows(anova(mod_2pl, mod_cfa),
          anova(mod_cfa, mod_bif)[2,], # 2pl 1d (and 2d) preferred
          anova(mod_cfa, mod_2pl_2f)[2,]) # AIC prefers bifactor, BIC favors CFA
```
The less conservative AIC prefers 3-factor over 2-factor or unidimensional models, but BIC prefers the simpler models. 
AIC prefers the bifactor model, while BIC prefers the CFA -- but the unidimensional 2PL is preferred by both criteria. 
