# Data loading

```{r}
library(tidyverse)
library(glue)
library(here)
library(mirt)
library(ggrepel)
# require(arm)
library(lavaan)
library(tidySEM)

source(here("02_score_data","irt_helpers.R"))
source(here("03_summaries", "plotting_helper.R"))

sites <- c("ca_pilot", "co_pilot", "de_pilot")

task_data_nested <- sites |>
  set_names() |>
  map(\(s) read_rds(here(glue("01_processed_data/{s}/task_data_nested.rds")))) |>
  list_rbind(names_to = "site")

task_data_combined <- task_data_nested |>
  select(-task_id) |>
  unnest(data) 

ef <- filter(task_data_combined, 
             task_id %in% c("hearts-and-flowers","same-different-selection","memory-game"))
```

Get ages. 

```{r}
participants <- sites |>
  set_names() |>
  map(\(s) read_rds(here(glue("00_prepped_data/{s}/participants.rds")))) |>
  list_rbind(names_to = "site")

run_ages <- participants |>
  select(user_id, ages) |>
  unnest(ages)

# this is useful below for various simplifications
ages <- run_ages |>
  group_by(user_id) |>
  summarise(age = mean(age))

ef <- left_join(ef, run_ages)
```

# Load multigroup models


```{r}
multigroup_best <- readRDS(here("02_scored_data", "irt_outputs", "multigroup_best_outputs.rds")) |>
  mutate(coefs = map(mod, multigroup_coefs))
multigroup_scores <- readRDS(here("02_scored_data", "scores", "scores_multigroup.rds"))
```



# Sumscore approach

First plot sumscores.

```{r}
ef_runs <- ef |>
  group_by(site, user_id, run_id, task_id) |>
  summarise(correct = mean(correct), 
            age = mean(age))

ggplot(ef_runs, aes(x = age, y = correct)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "loess", lambda = 1) +
  facet_grid(site ~ task_id) + 
  ylim(0,1) 
```

# Hearts and Flowers

Zoom in on HF. 

```{r}
hf <- ef |>
  filter(task_id == "hearts-and-flowers") 
```

Look at all phases. 

```{r}
hf_blocks <- hf |>
  mutate(corpus_trial_type = fct_relevel(corpus_trial_type, "hearts", "flowers", "mixed")) |>
  group_by(site, user_id, run_id, task_id, corpus_trial_type) |>
  summarise(correct = mean(correct), 
            age = mean(age), 
            n = n())

ggplot(hf_blocks,
       aes(x = age, y = correct)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  geom_hline(yintercept = .5, lty = 2) + 
  ylim(0,1) + 
  facet_grid(site ~ corpus_trial_type)
```

F notes that there are a lot of younger kids BELOW chance. Let's see if they are missing data. 

```{r}
ggplot(hf_blocks, aes(x = n, y = correct)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_grid(site ~ corpus_trial_type)
```

NO - this is not because of missing data. Something else is happening. 

Averaged across everything. 

```{r}
hf_runs <- hf |>
  group_by(site, user_id, run_id, task_id) |>
  summarise(logit = coef(glm(correct ~ 1, family = "binomial"))[1],
            bayeslogit = coef(arm::bayesglm(correct ~ 1, family = "binomial", 
                                            prior.scale.for.intercept = .1))[1],
            correct_rt = mean(as.numeric(rt[correct]), na.rm=TRUE),
            correct = mean(correct), 
            age = mean(age))


ggplot(hf_runs,
       aes(x = age, y = correct)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  geom_hline(yintercept = .5, lty = 2) + 
  ylim(0,1) + 
  facet_grid(site ~ task_id)
```
Logit scoring. 

```{r}
ggplot(hf_runs,
       aes(x = age, y = logit)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  geom_hline(yintercept = 0, lty = 2) + 
  # ylim(0,1) + 
  facet_grid(site ~ task_id)
```

Bayesian logit. 

```{r}
ggplot(hf_runs,
       aes(x = age, y = bayeslogit)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, lty = 2) + 
  # ylim(0,1) + 
  facet_grid(site ~ task_id)
```

Histograms. 

```{r}
ggplot(filter(ef_runs, task_id == "hearts-and-flowers") |>
         mutate(age_group = cut(age, c(5, 7, 9, 11, 13, include.lowest=TRUE))) |>
         filter(!is.na(age_group)),
       aes(x = correct)) + 
  geom_histogram(binwidth = .25) +
  geom_vline(xintercept = .5, lty = 2) +
  scale_x_continuous(breaks = c(0,.25, .5, .75, 1)) + 
  facet_grid(site ~ age_group)
```

Reaction time. 

```{r}
ggplot(hf_runs,
       aes(x = age, y = correct_rt)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, lty = 2) + 
  # ylim(0,1) + 
  facet_grid(site ~ task_id)
```

```{r}
ggplot(hf, aes(x = as.numeric(rt))) + 
  geom_histogram() + 
  scale_x_log10(breaks = c(100,200, 500,1000,2000,3000,5000,10000)) + coord_flip()
```


## IRT for H&F

We now use the multigroup IRT models. 

```{r}
multigroup_scores_hf <- multigroup_scores |>
  filter(task_id == "hearts-and-flowers") |>
  select(site, task_id, user_id, run_id, metric_type, metric_value ) |>
  left_join(run_ages)

ggplot(multigroup_scores_hf, aes(x = age, y = metric_value, col = site)) + 
  geom_point() + 
  geom_smooth()
```
Look at coefficients. 

```{r}
hf_coefs <- filter(multigroup_best, task_id == "hearts-and-flowers")$coefs[[1]] |>
  separate(item, sep = "_", into = c("task","block","trial", "number")) |>
  select(-number) |>
  distinct()

knitr::kable(hf_coefs, digits = 2)
```


# Memory game

```{r}
mg <- ef |>
  filter(task_id == "memory-game", 
         corpus_trial_type != "") |>
  mutate(span = as.numeric(item_id)) |>
  group_by(run_id) |>
  mutate(max_span = max(span))
```


```{r}
mg_spans <- mg |>
  group_by(site, span, corpus_trial_type) |>
  summarise(n = n(), 
            correct = mean(correct))

ggplot(mg_spans, aes(x = span, y = correct, col = corpus_trial_type)) + 
  geom_point(aes(size = n)) + 
  geom_line() + 
  facet_wrap(~site)
```

```{r}

mg |>
  filter(corpus_trial_type == "forward", span > 1) |>
  mutate(age_rounded = as_factor(round(age))) |>
  group_by(age_rounded, span) |>
  summarise(correct = mean(correct), 
            n = n()) |>
  ggplot(aes(x = span, y = correct, col = age_rounded)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(method = "lm", se =FALSE)

```

```{r}

mg |>
  filter(corpus_trial_type == "forward", span > 1) |>
  mutate(max_span = as_factor(max_span)) |>
  group_by(max_span, span) |>
  summarise(correct = mean(correct), 
            n = n()) |>
  ggplot(aes(x = span, y = correct, col = max_span)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(method = "lm", se =FALSE)

```



```{r}
mg_runs <- mg |>
  filter(correct) |>
  group_by(site, user_id, run_id, corpus_trial_type) |>
  summarise(max_span = max(span), 
            age = mean(age))
```


```{r}
ggplot(mg_runs, aes(x = age, y = max_span, col = corpus_trial_type)) + 
  geom_jitter(alpha = .5, height = .1, width = 0) + 
  geom_smooth(method = "lm")+ 
  facet_grid(corpus_trial_type~site, scales = "free_y")
```

## memory IRT


```{r}
mg_data_prepped <- mg |>
  ungroup() |>
  # filter(corpus_trial_type == "forward") |>
  mutate(item_id = str_c(corpus_trial_type, "_", as.character(span))) |>
  nest() |>
  mutate(data_filtered = map(data, \(df) df |> filter_repeat_runs() |>
                               dedupe_items() |> remove_no_var_items()),
         data_prepped = map(data_filtered, to_mirt_shape)) |>
  # pull out chance values
  mutate(guess = map(data_filtered, # TODO: check that this gives correct order
                     \(df) df |> distinct(item_inst, chance) |> pull(chance)))

# generate model strings, using constraints for repeated items

item_types <- c("Rasch", "2PL") #, "3PL") # set of parameterizations
model_types <- c(1) #, 2) # set of dimensionalities
# model_types <- c(1) # set of dimensionalities

# add arguments for model fitting to data
mg_data_args <- mg_data_prepped |>
  # duplicate rows per dimensionality x parameterization
  expand_grid(model_type = model_types, item_type = item_types) |>
  # generate model string with item constraints + dimensionality
  mutate(model_str = pmap_chr(list(data, data_prepped, item_type, model_type),
                              generate_model_str))

mg_models <- mg_data_args |>
  mutate(mod = pmap(list(row_number(), data_prepped, item_type, model_str,
                         model_type, guess),
                    fit_mirt))
```

```{r}
# get each model's coefs, scores, BIC
mg_results <- mg_models |>
  mutate(coefs = map(mod, mirt_coefs),
         scores = pmap(list(mod, data_filtered, data_prepped), mirt_scores),
         bic = map_dbl(mod, mirt_bic))

# best fitting model for each task
mg_best <- mg_results |>
  filter(bic == min(bic)) |>
  ungroup()
# select(site, task_id, item_type, model_type, coefs, scores)

# scores from best fitting models
mg_scores <- mg_best |>
  select(item_type, model_type, scores) |>
  unnest(scores) |>
  mutate(#item_type = fct_recode(item_type, "1PL" = "Rasch"),
    metric_type = glue("ability ({item_type}-{model_type})")) |>
  select(user_id, run_id, metric_type, metric_value = ability)

```

```{r}
mg_coefs <- mg_best$coefs[[1]] |>
  mutate(item = fct_reorder(item, d)) |>
  separate(item, sep = "_", into = c("direction","span","number"))
```

```{r}
ggplot(mg_coefs, 
       aes(x = span, y = -d, col = direction)) + 
  geom_point() + 
  geom_smooth(aes(group = direction), method = "lm") + 
  ylab("difficulty")
```

```{r}
ggplot(left_join(mg_best$scores[[1]], ages), 
       aes(x = age, y = ability)) +
  geom_point() + 
  geom_smooth(method = "lm")
```


```{r}
mg_runs_scores <- left_join(mg_runs, mg_best$scores[[1]]) |>
  mutate(task_id = "memory-game") |>
  filter(corpus_trial_type == "forward") |>
  mutate(rasch = ability) 

ggplot(mg_runs_scores,
       aes(x = max_span, y = ability)) +
  geom_point() + 
  geom_smooth(method = "lm")
```


# SDS

```{r}
sds <- filter(ef, task_id == "same-different-selection")

sds_runs <- sds |>
  group_by(site, user_id, run_id, task_id) |>
  summarise(correct = mean(correct), 
            age = mean(age))

sds_runs_block <- sds |>
  group_by(site, user_id, run_id, task_id, corpus_trial_type) |>
  summarise(correct = mean(correct), 
            age = mean(age))


ggplot(sds_runs_block, aes(x = age, y = correct)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "loess", lambda = 1) + 
  facet_wrap(~corpus_trial_type)
```


```{r}
parse_match_ids <- function (ii) {
 sapply(str_split(ii," "), 
      \(str) paste(str[1], str[3]))
}


sds_data_prepped <- sds |>
  mutate(item_id = case_when(
    str_detect(item_id, "test_dimensions") ~ str_sub(item_id, 1, -3),
    str_detect(item_id, "something_same_2") ~ str_sub(item_id, 1, -3),
    str_detect(item_id, "match") ~ parse_match_ids(item_id),
    .default = NA)) |>
  mutate(item_id = str_replace_all(item_id, " ", "_")) |>
  ungroup() |>
  # filter(corpus_trial_type == "forward") |>
  # mutate(item_id = str_c(corpus_trial_type, "_", as.character(span))) |>
  nest() |>
  mutate(data_filtered = map(data, \(df) df |> filter_repeat_runs() |>
                               dedupe_items() |> remove_no_var_items()),
         data_prepped = map(data_filtered, to_mirt_shape)) |>
  # pull out chance values
  mutate(guess = map(data_filtered, # TODO: check that this gives correct order
                     \(df) df |> distinct(item_inst, chance) |> pull(chance)))

# generate model strings, using constraints for repeated items

item_types <- c("Rasch", "2PL") #, "3PL") # set of parameterizations
model_types <- c(1) #, 2) # set of dimensionalities
# model_types <- c(1) # set of dimensionalities

# add arguments for model fitting to data
sds_data_args <- sds_data_prepped |>
  # duplicate rows per dimensionality x parameterization
  expand_grid(model_type = model_types, item_type = item_types) |>
  # generate model string with item constraints + dimensionality
  mutate(model_str = pmap_chr(list(data, data_prepped, item_type, model_type),
                              generate_model_str))

sds_models <- sds_data_args |>
  mutate(mod = pmap(list(row_number(), data_prepped, item_type, model_str,
                         model_type, guess),
                    fit_mirt))
```

```{r}
# get each model's coefs, scores, BIC
sds_results <- sds_models |>
  mutate(coefs = map(mod, mirt_coefs),
         scores = pmap(list(mod, data_filtered, data_prepped), mirt_scores),
         bic = map_dbl(mod, mirt_bic))

# best fitting model for each task
sds_best <- sds_results |>
  filter(bic == min(bic)) |>
  ungroup()
# select(site, task_id, item_type, model_type, coefs, scores)

# scores from best fitting models
sds_scores <- sds_best |>
  select(item_type, model_type, scores) |>
  unnest(scores) |>
  mutate(#item_type = fct_recode(item_type, "1PL" = "Rasch"),
    metric_type = glue("ability ({item_type}-{model_type})")) |>
  select(user_id, run_id, metric_type, metric_value = ability)

```

```{r}
sds_coefs <- sds_best$coefs[[1]] |>
  mutate(item = fct_reorder(item, d)) 
```


```{r}
sds_runs_scores <- left_join(sds_scores, sds_runs)
ggplot(sds_runs_scores, aes(x = age, y = metric_value)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~site)
```




# EF factor analysis

```{r}
mg_scores_cleaned <- mg_runs_scores |> 
  filter(corpus_trial_type == "forward") |>
  pivot_longer(c(max_span,ability), 
               names_to = "metric_type", values_to = "metric_value") |>
  select(site, task_id, user_id, run_id, metric_type,  metric_value)

sds_scores_cleaned <- sds_runs_scores |> 
  select(site, task_id, user_id, run_id, metric_value, correct) |>
  rename(ability = metric_value) |>
  pivot_longer(c(correct, ability), names_to = "metric_type", values_to = "metric_value")

hf_scores_cleaned <- bind_rows(hf_runs_score |>
              pivot_longer(c(correct, logit, bayeslogit, two_pl), names_to = "metric_type", 
                           values_to = "metric_value") |>
              select(site, task_id, user_id, run_id, metric_type, metric_value), 
              hf_scores |>
              left_join(select(participants, site, user_id)) |>
              mutate(task_id = "hearts-and-flowers"))

ef_scores <- bind_rows(mg_scores_cleaned, sds_scores_cleaned) |>
  bind_rows(hf_scores_cleaned) 
```

```{r}
ungroup(ef_scores) |>
  select(task_id, metric_type) |> 
  distinct()
```


```{r}
ef_scores_sumscore <- ef_scores |>
  filter(metric_type %in% c("max_span","correct")) |>
  group_by(site, task_id, user_id) |>
  summarise(metric_value = mean(metric_value, na.rm=TRUE)) |>
  ungroup() |>
  select(site, task_id, metric_value, user_id) |>
  pivot_wider(names_from = "task_id", values_from = "metric_value",
              id_cols = c("user_id","site")) |>
  janitor:::clean_names() |>
  mutate(across(same_different_selection:hearts_and_flowers, ~ scale(.x)[,1])) |>
  left_join(ages)

ef_scores_irt <- ef_scores |>
  filter(str_detect(metric_type, "ability")) |>
  group_by(site, task_id, user_id) |>
  summarise(metric_value = mean(metric_value, na.rm=TRUE)) |>
  ungroup() |>
  select(site, task_id, metric_value, user_id) |>
  pivot_wider(names_from = "task_id", values_from = "metric_value", 
              id_cols = c("user_id", "site")) |>
  janitor:::clean_names() |>
  mutate(across(same_different_selection:hearts_and_flowers, ~ scale(.x)[,1])) |>
  left_join(ages)
```

```{r}
cfa_model <-  "
ef =~ hearts_and_flowers + memory_game + same_different_selection
ef ~ age
"

fit <- cfa(cfa_model, ef_scores_sumscore, std.lv=TRUE, missing='fiml')
summary(fit, fit.measures=TRUE, standardize=TRUE)

graph_sem(model = fit, text_size = 3) + 
  theme(panel.background = element_rect(fill = "white"))
```


```{r}
cfa_model <-  "
ef =~ hearts_and_flowers + memory_game + same_different_selection
ef ~ age
"

fit_irt <- cfa(cfa_model, ef_scores_irt, std.lv=TRUE, missing='fiml')
summary(fit_irt, fit.measures=TRUE, standardize=TRUE)

graph_sem(model = fit_irt, text_size = 3) + 
  theme(panel.background = element_rect(fill = "white"))
```
