---
format: html
---

```{r setup}
source("_setup.R")

trials_coded <- read_rds("data_processed/trials_coded.rds")
participants <- read_rds("data_processed/participants.rds") |>
  # Get school name (assume ColegB = ColB)
  mutate(school = group_name |> gsub(pattern="Coleg", replacement="Col") |> str_extract("Col.") |> substring(4))

# Number of participants missing group name
nrow(participants |> filter(is.na(group_name))) # 367 (down from 1400)
```

# Load all scores
```{r}
score_list <- map(list.files("scores", full.names = TRUE), read_rds)

task_categories <- tribble(
  ~task, ~task_category,
  "hearts-and-flowers", "executive function",
  "same-different-selection", "executive function",
  "memory-game", "executive function",
  "egma-math", "math",
  "matrix-reasoning", "reasoning",
  "mental-rotation", "spatial cognition",
  "trog", "language",
  "vocab", "language",
  "pa-es", "reading",
  "sre-es", "reading",
  "swr-es", "reading",
  "theory-of-mind", "social cognition",
  "hostile-attribution", "social cognition"
) |> mutate(task_category = task_category |> str_to_sentence() |> fct_inorder())


scores <- score_list |>
  bind_rows() |> 
  distinct() |> # Remove duplicate rows from IRT task scores
  left_join(participants) |>
  filter(!is.na(school), task != "hostile-attribution") |>
  # filter(!is.na(age), age >= 5, age <= 12) |> # Missing many ages
  group_by(task) |>
  mutate(n = n_distinct(user_id), task_label = glue("{task}\n(n = {n})")) |>
  ungroup() |>
  left_join(task_categories) |>
  mutate(score_type = score_type |> str_remove(" \\(.*\\)"))

task_cat <- levels(scores$task_category)
task_pal <- ptol_pal()(length(task_cat)) |> set_names(task_cat)
```

Scores with missing school info
```{r}
scores_no_school <- score_list |>
  bind_rows() |> 
  distinct() |> # Remove duplicate rows from IRT task scores
  left_join(participants) |>
  filter(is.na(school), task != "hostile-attribution")

# Entries with missing school info for each task
#    task                     `n()`
#    <chr>                    <int>
#  1 egma-math                   38
#  2 hearts-and-flowers         161
#  3 matrix-reasoning            20
#  4 memory-game                145
#  5 mental-rotation             23
#  6 same-different-selection   168
#  7 sre-es                     204
#  8 swr-es                     204
#  9 theory-of-mind              23
# 10 trog                        26
# 11 vocab                       71
scores_no_school |> group_by(task) |> summarise(n())
nrow(scores_no_school) # 1083

# Number of participants with scores but missing school info
length(unique(scores_no_school$user_id)) # 625
```

# Distributions across all schools

```{r}
dist_all_schools <- scores |> 
  group_by(task_category, task) |>
  summarise(n = n(), Mean = mean(score), 
            Median = median(score), SD = sd(score)) |>
  rename(c("Category"="task_category", "Task"="task"))

dist_all_schools
```

All scores combining all schools in one plot
```{r}
ggplot(scores, aes(x = "", y = score)) +
  ggh4x::facet_nested_wrap(vars(task_category, task_label), nrow = 2,
                           nest_line = element_line(), solo_line = TRUE,
                           axes = "x",scales = "free_y") +
  geom_sina(aes(colour = task_category), alpha = 0.8) +
  scale_colour_manual(values = task_pal) +
  labs(y = "Score") +
  guides(colour = "none") +
  theme(axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10))

ggsave("plots/reports/all_scores_all_schools.png", width = 14, height = 8)
```

Just math and reading
```{r}
ggplot(scores |> filter(task_category %in% c("Math", "Reading")), 
       aes(x = "", y = score)) +
  ggh4x::facet_nested_wrap(vars(task_category, task_label), nrow = 1,
                           nest_line = element_line(), solo_line = TRUE,
                           axes = "x",scales = "free_y") +
  geom_sina(aes(colour = task_category), alpha = 0.8) +
  scale_colour_manual(values = task_pal) +
  labs(y = "Score") +
  guides(colour = "none") +
  theme(axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 15))
ggsave("plots/reports/math_read_all_schools.png", width = 14, height = 8)
```

# Distributions by school

```{r}
dist_by_school <- scores |> 
  group_by(school, task_category, task) |>
  summarise(n = n(), Mean = mean(score), 
            Median = median(score), SD = sd(score)) |>
  rename(c("Category"="task_category", "Task"="task"))

dist_by_school
```

All scores, sina plots
```{r}
ggplot(scores, aes(x = school, y = score)) +
  ggh4x::facet_nested_wrap(vars(task_category, task_label), nrow = 2,
                           nest_line = element_line(), solo_line = TRUE,
                           axes = "x",scales = "free_y") +
  geom_sina(aes(color = task_category), alpha = 0.8) + 
  scale_color_manual(values = task_pal) +
  labs(x="School", y = "Score") +
  guides(color = "none") +
  theme(strip.text = element_text(size = 10))

ggsave("plots/reports/all_scores_by_school_sina.png", width = 14, height = 8)
```

All scores, violin plots
```{r}
ggplot(scores, aes(x = school, y = score)) +
  ggh4x::facet_nested_wrap(vars(task_category, task_label), nrow = 2,
                           nest_line = element_line(), solo_line = TRUE,
                           axes = "x",scales = "free_y") +
  geom_violin(trim = F, aes(fill = task_category), alpha = 0.8) + 
  geom_boxplot(width=0.1) +
  scale_fill_manual(values = task_pal) +
  labs(x="School", y = "Score") +
  guides(fill = "none") +
  theme(strip.text = element_text(size = 10))

ggsave("plots/reports/all_scores_by_school_violin.png", width = 14, height = 8)
```

All scores, one school on one plot
```{r}
school_plot <- function(sch) {
  scores |> filter(school==sch) |> group_by(task) |>
    mutate(n = n_distinct(user_id), task_label = glue("{task}\n(n = {n})")) |>
    ungroup() |>
    ggplot(aes(x = "", y = score)) +
    ggh4x::facet_nested_wrap(vars(task_category, task_label), nrow = 2,
                           nest_line = element_line(), solo_line = TRUE,
                           axes = "x",scales = "free_y") +
    geom_sina(aes(colour = task_category), alpha = 0.8) +
    scale_colour_manual(values = task_pal) +
    labs(y = "Score") +
    guides(colour = "none") +
    theme(axis.ticks.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_text(size = 10))
}

school_plot("A")
ggsave("plots/reports/all_scores_schoolA.png", width = 14, height = 8)

school_plot("B")
ggsave("plots/reports/all_scores_schoolB.png", width = 14, height = 8)

school_plot("C")
ggsave("plots/reports/all_scores_schoolC.png", width = 14, height = 8)

school_plot("D")
ggsave("plots/reports/all_scores_schoolD.png", width = 14, height = 8)
```

Just math & reading scores, one school on one plot
```{r}
school_math_read_plot <- function(sch) {
  scores |> filter(school==sch, task_category %in% c("Math", "Reading")) |> 
    group_by(task) |> 
    mutate(n = n_distinct(user_id), task_label = glue("{task}\n(n = {n})")) |>
    ungroup() |>
    ggplot(aes(x = "", y = score)) +
    ggh4x::facet_nested_wrap(vars(task_category, task_label), nrow = 1,
                             nest_line = element_line(), solo_line = TRUE,
                             axes = "x",scales = "free_y") +
    geom_sina(aes(colour = task_category), alpha = 0.8) +
    scale_colour_manual(values = task_pal) +
    labs(y = "Score") +
    guides(colour = "none") +
    theme(axis.ticks.x = element_blank(),
          axis.title.x = element_blank(),
          strip.text = element_text(size = 10))
}

school_math_read_plot("A")
ggsave("plots/reports/math_read_schoolA.png", width = 14, height = 8)

school_math_read_plot("B")
ggsave("plots/reports/math_read_schoolB.png", width = 14, height = 8)

school_math_read_plot("C")
ggsave("plots/reports/math_read_schoolC.png", width = 14, height = 8)

school_math_read_plot("D")
ggsave("plots/reports/math_read_schoolD.png", width = 14, height = 8)
```

# Save distribution files

Combine all distributions into one csv
```{r}
dist_all_schools <- dist_all_schools |> mutate(school = "all")
all_dist <- rbind(dist_all_schools, dist_by_school)
write_csv(all_dist, "data/scores_summary_reports.csv")
```

Top 5% in each task from each school
```{r}
top_5_percent <- function(df) {
  threshold <- quantile(df$score, 0.95)
  return(df |> filter(score >= threshold))
}

top_5_percent_users <- scores |>
  group_by(school, task) |>
  do(top_5_percent(.)) |>
  select(school, task_category, task, user_id, score)
write_csv(top_5_percent_users, "data/top5perc_task_school.csv")
```

Bottom 5% in each task from each school
```{r}
bottom_5_percent <- function(df) {
  threshold <- quantile(df$score, 0.05)
  return(df |> filter(score <= threshold))
}

bottom_5_percent_users <- scores |>
  group_by(school, task) |>
  do(bottom_5_percent(.)) |>
  select(school, task_category, task, user_id, score)

write_csv(bottom_5_percent_users, "data/bottom5perc_task_school.csv")
```

