---
title: "Spark_h&f"
format: html
editor: visual
---
# Applying the Levante IRT Model to the PLUS and Kimochis datasets
The goal of this analysis is to evaluate the Levante Hearts and Flowers (H&F) IRT model's performance on an external dataset collected by Spark Lab, as a test of the model's generalizability. The Spark Lab data consists of two studies: PLUS (3rd–5th graders, group-based tablet assessment) and Kimochis (preschoolers, 1:1 assessment). 
```{r}
#| label: packages
#| include: false
library(tidyverse)   
library(mirt)        
library(here)
library(glue)
library(flextable)
library(patchwork)
library(janitor)
library(haven)

```

```{r}
#| label: load-data
#| include: false

plus     <- readRDS("plus_hf_merged.rds") |> mutate(pl_fem = as_factor(pl_fem))
kimochis <- readRDS("kimochis_hf.rds")
```

```{r}
#| label: explore-data
#| include: false

head(plus, 3)
head(kimochis, 3)

unique(plus$block)
unique(plus$stim_shape)

unique(kimochis$block)
unique(kimochis$stim_shape)

# check structure
glimpse(plus)
glimpse(kimochis)

# ccount of what blocks are present
plus |> count(block) 
kimochis |> count(block)

# check shape of PLUS
plus |> 
  filter(block == "flowers test") |>
  count(stim_shape, resp_side, stim_side) |>
  head(20)
```

## Part 1: Sample Descriptives
```{r}
#| label: sample-descriptives-table
#| echo: false

plus_sample <- plus |>
  group_by(time) |>
  summarise(
    dataset    = "PLUS",
    n_children = n_distinct(stud_id),
    age_mean   = mean(pl_age, na.rm = TRUE),
    age_sd     = sd(pl_age,  na.rm = TRUE),
    pct_female = mean(pl_fem == "Female", na.rm = TRUE) * 100,
    .groups    = "drop"
  ) |>
  mutate(time = as.character(time))

kimo_sample <- kimochis |>
  group_by(time) |>
  summarise(
    dataset    = "Kimochis",
    n_children = n_distinct(id),
    age_mean   = mean(age, na.rm = TRUE),
    age_sd     = sd(age,  na.rm = TRUE),
    pct_female = mean(female, na.rm = TRUE) * 100,
    .groups    = "drop"
  ) |>
  mutate(time = as.character(time))

plus_retention <- plus |>
  group_by(stud_id) |>
  summarise(n_tp = n_distinct(time)) |>
  count(n_tp) |>
  mutate(dataset = "PLUS", label = glue("{n_tp} tp: n = {n}")) |>
  group_by(dataset) |>
  summarise(retention = paste(label, collapse = "\n"), .groups = "drop")

kimo_retention <- kimochis |>
  group_by(id) |>
  summarise(n_tp = n_distinct(time)) |>
  count(n_tp) |>
  mutate(dataset = "Kimochis", label = glue("{n_tp} tp: n = {n}")) |>
  group_by(dataset) |>
  summarise(retention = paste(label, collapse = "\n"), .groups = "drop")

bind_rows(plus_sample, kimo_sample) |>
  left_join(bind_rows(plus_retention, kimo_retention), by = "dataset") |>
  mutate(
    age_m_sd   = glue("{round(age_mean, 2)} ({round(age_sd, 2)})"),
    pct_female = round(pct_female, 1)
  ) |>
  select(dataset, time, n_children, age_m_sd, pct_female, retention) |>
  rename(
    Dataset      = dataset,
    Timepoint    = time,
    N            = n_children,
    `Age M (SD)` = age_m_sd,
    `% Female`   = pct_female,
    Retention    = retention
  ) |>
  flextable() |>
  colformat_double(digits = 1) |>
  merge_v(j = c("Dataset", "Retention")) |>
  valign(j = c("Dataset", "Retention"), valign = "top") |>
  set_table_properties(width = 1, layout = "autofit") |>
  width(j = "Retention", width = 1.8) |>
  width(j = "Age M (SD)", width = 1.2) |>
  align(j = c("N", "% Female"), align = "right", part = "all") |>
  bold(part = "header") |>
  fontsize(size = 11, part = "all") |>
  theme_vanilla()
```

```{r}
#| label: kimochis-age-check
#| include: false

# are the exact same children appearing across timepoints?
kimochis |>
  group_by(id) |>
  summarise(n_timepoints = n_distinct(time)) |>
  count(n_timepoints)

# do the min/max age children appear at multiple timepoints with the same age?
kimochis |>
  filter(age == min(age, na.rm = TRUE) | age == max(age, na.rm = TRUE)) |>
  distinct(id, time, age, grade) |>
  arrange(id, time)

# is age time-varying within a child, or fixed at enrolment?
kimochis |>
  group_by(id) |>
  summarise(n_distinct_ages = n_distinct(age)) |>
  count(n_distinct_ages)

# distribution of age at each timepoint
kimochis |>
  ggplot(aes(x = age, fill = time)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(~time) +
  theme_minimal() +
  labs(title = "Age distribution by timepoint")

```

```{r}
#| label: sample-reporting-plus
#| include: false

plus_n_total <- n_distinct(plus$stud_id)
plus_n_schools <- n_distinct(plus$sch_id)

plus_summary <- plus |>
  group_by(time) |>
  summarise(
    n = n_distinct(stud_id),
    age_mean = mean(pl_age, na.rm = TRUE) |> round(2),
    age_sd   = sd(pl_age,  na.rm = TRUE)  |> round(2),
    .groups  = "drop"
  )

plus_n_fall   <- plus_summary |> filter(time == "PLUS Fall")   |> pull(n)
plus_n_spring <- plus_summary |> filter(time == "PLUS Spring") |> pull(n)
plus_age_mean <- plus_summary |> pull(age_mean) |> mean() |> round(2)
plus_age_sd   <- plus_summary |> pull(age_sd)   |> mean() |> round(2)
```
### Plus
The Plus dataset includes `r plus_n_total` unique students across two timepoints (PLUS Fall and PLUS Spring) drawn from `r plus_n_schools` schools. Sample sizes were stable across timepoints (Fall: n = `r plus_n_fall`, Spring: n = `r plus_n_spring`). Mean age was approximately `r plus_age_mean` years (SD ~`r plus_age_sd`) at both timepoints, with students in grades 3, 4, and 5 roughly evenly distributed. The sample was approximately 49% female at both timepoints.

```{r}
#| label: sample-reporting-kimochis
#| include: false

kimo_n_total <- n_distinct(kimochis$id)
kimo_n_t1 <- kimochis |> filter(time == "2021-2022 Fall")   |> summarise(n = n_distinct(id)) |> pull(n)
kimo_n_t4 <- kimochis |> filter(time == "2022-2023 Spring") |> summarise(n = n_distinct(id)) |> pull(n)

# from the sample descriptives summary table
kimo_summary <- kimochis |>
  group_by(time) |>
  summarise(
    n        = n_distinct(id),
    age_mean = mean(age, na.rm = TRUE) |> round(2),
    age_sd   = sd(age,  na.rm = TRUE)  |> round(2),
    .groups  = "drop"
  )

kimo_age_min <- kimo_summary |> summarise(m = min(age_mean)) |> pull(m)
kimo_age_max <- kimo_summary |> summarise(m = max(age_mean)) |> pull(m)

kimo_retention <- kimochis |>
  group_by(id) |>
  summarise(n_timepoints = n_distinct(time)) |>
  count(n_timepoints)

kimo_n_1tp  <- kimo_retention |> filter(n_timepoints == 1) |> pull(n)
kimo_n_2tp  <- kimo_retention |> filter(n_timepoints == 2) |> pull(n)
kimo_n_3tp  <- kimo_retention |> filter(n_timepoints == 3) |> pull(n)
kimo_n_4tp  <- kimo_retention |> filter(n_timepoints == 4) |> pull(n)

kimo_n_fixed_age <- kimochis |>
  group_by(id) |>
  summarise(n_distinct_ages = n_distinct(age)) |>
  filter(n_distinct_ages == 1) |>
  nrow()
```
### Kimochis
The Kimochis dataset includes `r kimo_n_total` unique children across four timepoints spanning two academic years (2021-2022 and 2022-2023, Fall and Spring). Sample sizes ranged from `r kimo_n_t1` children at the first timepoint to `r kimo_n_t4` at the last. The sample was approximately evenly split by sex (~46-51% female) across timepoints, with mean age ranging from `r kimo_age_min` to `r kimo_age_max` years. Children were enrolled in Pre-K and, from 2021-2022 Spring onwards, Transitional Kindergarten (TK).

Longitudinal retention was limited: the majority of children appeared at only one timepoint
(n = `r kimo_n_1tp`), with `r kimo_n_2tp` appearing at two, `r kimo_n_3tp` at three, and 
`r kimo_n_4tp` at all four timepoints. The dataset is therefore better characterised as largely cross-sectional with limited repeated measures.

Age in the Kimochis dataset should be interpreted with caution. For the majority of children
(n = `r kimo_n_fixed_age`), age did not vary across timepoints, suggesting it reflects 
approximate age at enrolment rather than age at time of assessment. Age values also appear to be rounded to the nearest month rather than recorded as a continuous measure. These features should be considered when using age as a covariate in models.

## Trial-level descriptive statistics by dataset and block
```{r}
#| label: descriptive-stats
#| include: false

# plus cleaning
plus <- plus |>
  mutate(
    dataset    = "plus",
    age        = pl_age,
    pl_fem     = as_factor(pl_fem),
    stim_shape = str_to_lower(stim_shape),
    stim_side  = str_to_lower(stim_side),
    resp_side  = str_to_lower(resp_side),
    acc        = case_when(
      stim_shape == "heart"  ~ as.numeric(resp_side == stim_side),
      stim_shape == "flower" ~ as.numeric(resp_side != stim_side),
      TRUE                   ~ NA_real_
    ),
    acc = if_else(is.na(resp_side), 0, acc)
  ) |>
  filter(
    !block %in% c("hearts practice", "flowers practice"),
    rt > 0,
    !is.na(rt)
  )

# kimochis cleaning
kimochis <- kimochis |>
  mutate(
    dataset    = "kimochis",
    stim_shape = str_to_lower(stim_shape),
    stim_side  = str_to_lower(stim_side),
    resp_side  = str_to_lower(resp_side),
    rt         = rt_all * 1000
  ) |>
  filter(!block %in% c("hearts practice", "flowers practice"))

both_descriptive <- bind_rows(
  plus     |> mutate(time = as.character(time)),
  kimochis |> mutate(time = as.character(time))
)

# structure
glimpse(plus)
glimpse(kimochis)

# numeric summary
plus     |> select(where(is.numeric)) |> summary()
kimochis |> select(where(is.numeric)) |> summary()

# trial count by block
both_descriptive |>
  count(dataset, block) |>
  pivot_wider(names_from = dataset, values_from = n)

# accuracy by block
both_descriptive |>
  group_by(dataset, block) |>
  summarise(
    n        = n(),
    acc_mean = mean(acc, na.rm = TRUE),
    acc_sd   = sd(acc,  na.rm = TRUE),
    .groups  = "drop"
  )

```

```{r}
#| label: descriptive-table
#| echo: false

both_descriptive |>
  filter(!is.na(rt)) |>
  group_by(dataset, block) |>
  summarise(
    n_trials = n(),
    age_mean = mean(age, na.rm = TRUE),
    age_sd   = sd(age,  na.rm = TRUE),
    acc_mean = mean(acc, na.rm = TRUE),
    acc_sd   = sd(acc,  na.rm = TRUE),
    rt_mean  = mean(rt),
    rt_med   = median(rt),
    rt_sd    = sd(rt),
    rt_min   = min(rt),
    rt_max   = max(rt),
    .groups  = "drop"
  ) |>
  mutate(age_m_sd = glue("{round(age_mean, 2)} ({round(age_sd, 2)})")) |>
  select(-age_mean, -age_sd) |>
  arrange(dataset, block) |>
  rename(
    Dataset     = dataset,
    Block       = block,
    `N Trials`  = n_trials,
    `Age M (SD)` = age_m_sd,
    `Acc Mean`  = acc_mean,
    `Acc SD`    = acc_sd,
    `RT Mean`   = rt_mean,
    `RT Median` = rt_med,
    `RT SD`     = rt_sd,
    `RT Min`    = rt_min,
    `RT Max`    = rt_max
  ) |>
  flextable() |>
  colformat_double(digits = 2) |>
  merge_v(j = "Dataset") |>
  theme_vanilla() |>
  autofit()

```

```{r}
#| label: plots
#| echo: false

# RT distributions by block
p_rt_dist <- both_descriptive |>
  filter(!is.na(rt)) |>
  ggplot(aes(x = rt, fill = dataset)) +
  geom_histogram(bins = 60, alpha = 0.7, position = "identity") +
  facet_wrap(~block, scales = "free_y") +
  labs(title = "RT distributions by block", x = "RT (ms)", y = "count") +
  theme_minimal()

# RT boxplots by block x stim_shape
p_rt_box <- both_descriptive |>
  filter(!is.na(rt)) |>
  ggplot(aes(x = block, y = rt, fill = stim_shape)) +
  geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
  facet_wrap(~dataset) +
  labs(title = "RT by block × stim_shape", y = "RT (ms)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

# accuracy by block x stim_shape
p_acc <- both_descriptive |>
  group_by(dataset, block, stim_shape) |>
  summarise(acc = mean(acc, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = block, y = acc, fill = stim_shape)) +
  geom_col(position = "dodge") +
  facet_wrap(~dataset) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Accuracy by block × stim_shape", y = "proportion correct") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

print(p_rt_dist)
print(p_rt_box)
print(p_acc)

```

## Part 2: Fixed Parameter Scoring
```{r}
#| label: data-quality-checks
#| include: false

# plus RT distribution checks
plus |>
  summarise(
    p95         = quantile(rt, 0.95),
    p99         = quantile(rt, 0.99),
    n_over_3000 = sum(rt > 3000),
    n_over_5000 = sum(rt > 5000)
  )

# kimochis: extreme RTs
kimochis |>
  filter(rt > 10000) |>
  select(id, block, trial_num, rt, acc, to) |>
  arrange(desc(rt))

# kimochis: timeout flag alignment 
# check whether to, rt NA, and acc NA are consistent
kimochis |>
  mutate(rt_na     = is.na(rt),
         acc_na    = is.na(acc),
         timed_out = to == 1) |>
  count(rt_na, acc_na, timed_out)

# plus: NA resp_side breakdown 
plus |>
  filter(is.na(resp_side)) |>
  count(block)

# kimochis: ft_stim NAs by block
kimochis |>
  filter(is.na(ft_stim)) |>
  count(block)

# kimochis: ar NAs vs acc NAs
kimochis |>
  mutate(ar_na  = is.na(ar),
         acc_na = is.na(acc)) |>
  count(ar_na, acc_na)

# missing values per column 
map_df(
  list(plus = plus, kimochis = kimochis),
  ~summarise(.x, across(everything(), ~sum(is.na(.)))),
  .id = "dataset"
)

# cell balance 
both_descriptive |>
  count(dataset, block, stim_shape, stim_side) |>
  group_by(dataset, block) |>
  mutate(expected  = mean(n),
         deviation = n - expected) |>
  filter(abs(deviation) > 5)

# duplicate rows
both_descriptive |>
  group_by(dataset) |>
  summarise(n_total    = n(),
            n_distinct = n_distinct(across(everything())),
            n_dupes    = n_total - n_distinct)

# congruency rule violations 
# for mixed test, rule is determined by stim_shape within trial
both_descriptive |>
  mutate(
    congruent          = stim_side == resp_side,
    expected_congruent = case_when(
      stim_shape == "heart"  ~ TRUE,
      stim_shape == "flower" ~ FALSE,
      TRUE                   ~ NA
    ),
    rule_violation = !is.na(expected_congruent) & (congruent != expected_congruent)
  ) |>
  count(dataset, block, rule_violation)

```

```{r}
#| label: levante-scores
#| include: false

# install remotes
# install.packages("remotes")
# install rlevante 
# remotes::install_github("levante-framework/rlevante")

best_coefs <- read_rds(here("02_scoring_outputs/best_coefs.rds"))
mods_best  <- read_rds(here("02_scoring_outputs/mods_best.rds"))

# inspect
glimpse(best_coefs)
glimpse(mods_best)
best_coefs |> distinct(task)

# get hf item parameters 
hf_coefs <- best_coefs |> filter(task == "hf")

# see all item names
hf_coefs |> distinct(item)

# pull the hf model record
hf_mod_rec <- mods_best |> filter(task == "hf") |> pull(mod_rec) |> pluck(1)

# inspect the model record slots
isVirtualClass("ModelRecord")
slotNames(hf_mod_rec)

# look at items slot specifically
hf_mod_rec@items
```

```{r}
#| label: levante-hf-structure
#| include: false

task_rxx <- read_rds(here("02_scoring_outputs/task_rxx.rds"))
colnames(task_rxx)
task_rxx |> filter(task == "hf")

# how many items in the levante hf model per site?
best_coefs |> 
  filter(task == "hf") |> 
  group_by(group) |> 
  distinct(item) |> 
  count()

# Load data
task_data_nested <- read_rds(here("01_fetched_data/task_data_nested.rds"))
glimpse(task_data_nested$data)

# look at the data structure for hf in levante
task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  group_by(site, user_id, run_id) |>
  summarise(n_trials = n(), .groups = "drop") |>
  group_by(site) |>
  summarise(mean_trials   = mean(n_trials),
            sd_trials     = sd(n_trials),
            median_trials = median(n_trials))

task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  group_by(site, user_id, run_id) |>
  summarise(n_unique_items = n_distinct(item_uid), .groups = "drop") |>
  group_by(site) |>
  summarise(mean_unique = mean(n_unique_items),
            median_unique = median(n_unique_items))
```

```{r}
#| label: prep-plus-hf-items
#| include: false

plus_hf_prepped <- plus |>
  filter(block %in% c("hearts test", "flowers test", "mixed test")) |>
  mutate(
    correct = case_when(
      stim_shape == "heart"  ~ resp_side == stim_side,
      stim_shape == "flower" ~ resp_side != stim_side,
      TRUE ~ NA
    ),
    correct     = if_else(is.na(resp_side), FALSE, correct),
    correct     = as.integer(correct),
    block_clean = case_when(
      block == "hearts test"  ~ "hearts",
      block == "flowers test" ~ "flowers",
      block == "mixed test"   ~ "heartsflowers"
    )
  )

# classify trials as start / stay / switch within each person x block
plus_hf_items <- plus_hf_prepped |>
  arrange(stud_id, time, block_clean, trial_num) |>
  group_by(stud_id, time, block_clean) |>
  mutate(
    prev_shape = lag(stim_shape),
    trial_type = case_when(
      trial_num == min(trial_num) ~ "start",
      stim_shape == prev_shape   ~ "stay",
      stim_shape != prev_shape   ~ "switch"
    )
  ) |>
  group_by(stud_id, time, block_clean, stim_shape, trial_type) |>
  mutate(trial_type_num = row_number()) |>
  ungroup() |>
  mutate(item = glue("{block_clean}_{stim_shape}_{trial_type}-{trial_type_num}"))

# timeout rate (flowers and mixed blocks only)
mean_timeout_pct <- plus_hf_items |>
  filter(block_clean != "hearts") |>
  group_by(stud_id, time) |>
  summarise(prop_timeout = mean(is.na(resp_side)), .groups = "drop") |>
  summarise(mean_timeout = round(mean(prop_timeout) * 100, 1)) |>
  pull(mean_timeout)

# attention check: exclude children scoring below 50% on hearts block
plus_hearts_check <- plus_hf_items |>
  filter(block_clean == "hearts") |>
  group_by(stud_id, time) |>
  summarise(hearts_acc = mean(correct, na.rm = TRUE), .groups = "drop") |>
  filter(hearts_acc >= 0.5)
```

```{r}
#| label: plus-hf-levante-equivalent
#| include: false

# collapse repeated trials of same type into single binary item
plus_hf_collapsed <- plus_hf_items |>
  semi_join(plus_hearts_check, by = c("stud_id", "time")) |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  group_by(stud_id, time, item_uid) |>
  summarise(prop_correct = mean(correct, na.rm = TRUE), .groups = "drop") |>
  mutate(correct_collapsed = as.integer(prop_correct >= 0.5))

# levante item_uid types
levante_uid_types <- c(
  "hearts_heart_start", "hearts_heart_stay",
  "flowers_flower_start", "flowers_flower_stay",
  "heartsflowers_heart_start", "heartsflowers_flower_start",
  "heartsflowers_heart_stay", "heartsflowers_flower_stay",
  "heartsflowers_heart_switch", "heartsflowers_flower_switch"
)

shared_uid_types <- intersect(
  plus_hf_collapsed |> distinct(item_uid) |> pull(item_uid),
  levante_uid_types
)

# pivot to wide and build response matrix
plus_hf_collapsed_wide <- plus_hf_collapsed |>
  select(stud_id, time, item_uid, correct_collapsed) |>
  pivot_wider(names_from = item_uid, values_from = correct_collapsed)

plus_collapsed_matrix <- plus_hf_collapsed_wide |>
  select(all_of(shared_uid_types)) |>
  as.matrix()
```

```{r}
#| label: plus-hf-levante-equivalent-rasch
#| include: false

# pull levante d parameters (identical across sites under scalar invariance)
levante_lookup_collapsed <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  filter(item_uid %in% levante_uid_types) |>
  group_by(item_uid, name) |>
  slice_min(item, n = 1) |>
  ungroup() |>
  distinct(item = item_uid, name, value)

# fix item parameters to levante values
fixed_pars_collapsed <- mirt(
  data     = plus_collapsed_matrix,
  model    = 1,
  itemtype = rep("Rasch", ncol(plus_collapsed_matrix)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

# fit model with fixed parameters and estimate EAP scores
hf_mod_plus_collapsed <- mirt(
  data      = plus_collapsed_matrix,
  model     = 1,
  itemtype  = rep("Rasch", ncol(plus_collapsed_matrix)),
  pars      = fixed_pars_collapsed,
  technical = list(NCYCLES = 1)
)

plus_scores_collapsed <- fscores(hf_mod_plus_collapsed, method = "EAP", full.scores.SE = TRUE)

plus_hf_scored_collapsed <- plus_hf_collapsed_wide |>
  select(stud_id, time) |>
  mutate(
    theta_collapsed = plus_scores_collapsed[, "F1"],
    se_collapsed    = plus_scores_collapsed[, "SE_F1"]
  )

plus_reliability_collapsed <- round(
  1 - mean(plus_hf_scored_collapsed$se_collapsed^2) / var(plus_hf_scored_collapsed$theta_collapsed),
  3
)

# grade-level theta summary
grade_summary <- plus_hf_scored_collapsed |>
  left_join(plus |> distinct(stud_id, pl_grade), by = "stud_id") |>
  group_by(pl_grade) |>
  summarise(
    m  = round(mean(theta_collapsed, na.rm = TRUE), 2),
    sd = round(sd(theta_collapsed, na.rm = TRUE), 2),
    .groups = "drop"
  )

g3_m  <- grade_summary |> filter(pl_grade == 3) |> pull(m)
g3_sd <- grade_summary |> filter(pl_grade == 3) |> pull(sd)
g4_m  <- grade_summary |> filter(pl_grade == 4) |> pull(m)
g4_sd <- grade_summary |> filter(pl_grade == 4) |> pull(sd)
g5_m  <- grade_summary |> filter(pl_grade == 5) |> pull(m)
g5_sd <- grade_summary |> filter(pl_grade == 5) |> pull(sd)
```
## Data preparation and scoring - PLUS 
Trial-level data from PLUS were filtered to test blocks only (hearts test, flowers test, mixed test), excluding practice trials. Accuracy was derived from the stimulus-response pairing: trials with a heart stimulus were scored correct if the response matched the stimulus side (congruent rule), and trials with a flower stimulus were scored correct if the response was on the opposite side (incongruent rule). Trials with no response (NA) were coded as incorrect. Children were excluded if they scored below 50% on the hearts block, treating this as an attention check consistent with standard Levante scoring practice. AAmong children retained after the attention check, `r mean_timeout_pct`% of trials in the flowers and mixed blocks had no response recorded and were coded as incorrect. This yielded a final analytic sample of `r n_distinct(plus_hf_scored_collapsed$stud_id)` children (`r format(nrow(plus_hf_scored_collapsed), big.mark = ",")` observations across fall and spring timepoints).

Each trial was assigned an item label following the Levante naming convention: `{block}_{stimulus}_{trial_type}-{trial_number}`, where trial type was classified as start (first trial in block), stay (same stimulus type as previous trial), or switch (different stimulus type from previous trial). Repeated trials of the same block-stimulus-trial_type combination were aggregated into a single analytic unit by taking the proportion correct across all trials of that type and binarizing at a 0.5 threshold. Item difficulty parameters were taken from the Levante H&F Rasch model, which was estimated under scalar invariance across sites; parameters are therefore identical across all three Levante sites. This yielded 9 IRT items in PLUS out of the 10 analytic units used in Levante scoring. The missing item type (heartsflowers_heart_start) was absent from PLUS due to differences in block structure between the two assessments.

Ability scores (θ) showed an increasing pattern across grades (3rd grade: M = `r g3_m`, SD = `r g3_sd`; 4th grade: M = `r g4_m`, SD = `r g4_sd`; 5th grade: M = `r g5_m`, SD = `r g5_sd`), consistent with the expected developmental trajectory and supporting the construct validity of the scoring approach in this sample. Grade-level means are close to the Levante scale origin and slightly positive, suggesting that 3rd–5th graders in PLUS perform at a broadly similar level to the Levante sample. 
```{r}
#| label: prep-kimochis-hf-items
#| include: false

# check stim_shape values (we saw these may be capitalized)
kimochis |> count(stim_shape)

# check acc vs acc_to 
kimochis |>
  filter(str_detect(block, "test")) |>
  summarise(
    mean_acc    = mean(acc, na.rm = TRUE),
    mean_acc_to = mean(acc_to, na.rm = TRUE),
    prop_timeout = mean(to, na.rm = TRUE)
  )

# check trial_num range per block
kimochis |>
  filter(str_detect(block, "test")) |>
  group_by(block) |>
  summarise(max_trial = max(trial_num), min_trial = min(trial_num))
####################

kimochis_hf_items <- kimochis |>
  filter(str_detect(block, "test")) |>
  mutate(
    stim_shape  = str_to_lower(stim_shape),
    block_clean = case_when(
      block == "hearts test"  ~ "hearts",
      block == "flowers test" ~ "flowers"
    )
  ) |>
  arrange(id, time, block_clean, trial_num) |>
  group_by(id, time, block_clean) |>
  mutate(
    # no switches within block — all stimuli are the same shape
    trial_type     = if_else(trial_num == min(trial_num), "start", "stay"),
    trial_type_num = if_else(trial_type == "start", 1L, row_number() - 1L)
  ) |>
  ungroup() |>
  mutate(item = paste0(block_clean, "_", stim_shape, "_", trial_type, "-", trial_type_num))

# exclude children scoring below 50% on hearts block
kimochis_hearts_check <- kimochis_hf_items |>
  filter(block_clean == "hearts") |>
  group_by(id, time) |>
  summarise(hearts_acc = mean(acc, na.rm = TRUE), .groups = "drop") |>
  filter(hearts_acc >= 0.5)
```

```{r}
#| label: kimochis-hf-levante-equivalent
#| include: false

# collapse repeated trials of same type into single binary item
kimochis_hf_collapsed <- kimochis_hf_items |>
  semi_join(kimochis_hearts_check, by = c("id", "time")) |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  group_by(id, time, item_uid) |>
  summarise(prop_correct = mean(acc, na.rm = TRUE), .groups = "drop") |>
  mutate(correct_collapsed = as.integer(prop_correct >= 0.5))

shared_uid_kimochis <- intersect(
  kimochis_hf_collapsed |> distinct(item_uid) |> pull(item_uid),
  levante_uid_types
)

# pivot to wide and build response matrix
kimochis_hf_collapsed_wide <- kimochis_hf_collapsed |>
  select(id, time, item_uid, correct_collapsed) |>
  pivot_wider(names_from = item_uid, values_from = correct_collapsed)

kimochis_collapsed_matrix <- kimochis_hf_collapsed_wide |>
  select(all_of(shared_uid_kimochis)) |>
  as.matrix()

```

```{r}
#| label: kimochis-hf-levante-equivalent-rasch
#| include: false

# fix item parameters to levante values
fixed_pars_collapsed_kimochis <- mirt(
  data     = kimochis_collapsed_matrix,
  model    = 1,
  itemtype = rep("Rasch", ncol(kimochis_collapsed_matrix)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

# fit model with fixed parameters and estimate EAP scores
hf_mod_kimochis_collapsed <- mirt(
  data      = kimochis_collapsed_matrix,
  model     = 1,
  itemtype  = rep("Rasch", ncol(kimochis_collapsed_matrix)),
  pars      = fixed_pars_collapsed_kimochis,
  technical = list(NCYCLES = 1)
)

kimochis_scores_collapsed <- fscores(hf_mod_kimochis_collapsed, method = "EAP", full.scores.SE = TRUE)

kimochis_hf_scored_collapsed <- kimochis_hf_collapsed_wide |>
  select(id, time) |>
  mutate(
    theta_collapsed = kimochis_scores_collapsed[, "F1"],
    se_collapsed    = kimochis_scores_collapsed[, "SE_F1"]
  )

kimochis_reliability_collapsed <- round(
  1 - mean(kimochis_hf_scored_collapsed$se_collapsed^2) / var(kimochis_hf_scored_collapsed$theta_collapsed),
  3
)

# age-group theta summary
kimochis_age_summary <- kimochis_hf_scored_collapsed |>
  left_join(kimochis |> distinct(id, time, age), by = c("id", "time")) |>
  mutate(age_group = cut(age, breaks = c(2, 3, 4, 5, 6),
                         labels = c("2-3", "3-4", "4-5", "5-6"))) |>
  group_by(age_group) |>
  summarise(
    m  = round(mean(theta_collapsed, na.rm = TRUE), 2),
    sd = round(sd(theta_collapsed, na.rm = TRUE), 2),
    .groups = "drop"
  )

a23_m  <- kimochis_age_summary |> filter(age_group == "2-3") |> pull(m)
a23_sd <- kimochis_age_summary |> filter(age_group == "2-3") |> pull(sd)
a34_m  <- kimochis_age_summary |> filter(age_group == "3-4") |> pull(m)
a34_sd <- kimochis_age_summary |> filter(age_group == "3-4") |> pull(sd)
a45_m  <- kimochis_age_summary |> filter(age_group == "4-5") |> pull(m)
a45_sd <- kimochis_age_summary |> filter(age_group == "4-5") |> pull(sd)
```

```{r}
#| label: sample-size-table
#| include: false

sample_sizes <- plus_hf_scored_collapsed |>
  count(time) |>
  mutate(sample = "PLUS", time = as.character(time)) |>
  bind_rows(
    kimochis_hf_scored_collapsed |>
      count(time) |>
      mutate(sample = "Kimochis", time = as.character(time))
  ) |>
  select(Sample = sample, Timepoint = time, N = n)

flextable(sample_sizes) |>
  merge_v(j = "Sample") |>
  theme_vanilla() |>
  autofit()
```
## Data preparation and scoring - Kimochis
Trial-level data from Kimochis were filtered to test blocks only (hearts test, flowers test), excluding practice trials. Unlike PLUS, the Kimochis dataset did not include a mixed block, as the children were considered too young for the task-switching demands it requires. Accuracy was taken from the pre-computed "acc" column rather than derived from stimulus-response pairings. Timeouts were treated as missing rather than incorrect, consistent with the 1:1 assessment context in which no strict response time limit was imposed. Children were excluded if they scored below 50% on the hearts block as an attention check. This yielded a final analytic sample of 1,622 children (2,540 observations) across two academic year cohorts (2021–2022 and 2022–2023), each with fall and spring timepoints.

Each trial was assigned an item label following the Levante naming convention. Of the 74 items in the Levante H&F model, 27 were present in Kimochis — all from the hearts and flowers blocks. The 8 additional flowers trials in Kimochis (stay-16 through stay-23) exceeded the length of the Levante flowers block and were excluded from scoring. All 46 mixed block items were absent by design. Repeated trials of the same block-stimulus-trial_type combination were aggregated into a single analytic unit by taking the proportion correct across all trials of that type and binarising at a 0.5 threshold, yielding 4 IRT items (hearts_heart_start, hearts_heart_stay, flowers_flower_start, flowers_flower_stay).

Ability scores (θ) showed an increasing pattern across age groups (2–3 years: M = `r a23_m`, SD = `r a23_sd`; 3–4 years: M = `r a34_m`, SD = `r a34_sd`; 4–5 years: M = `r a45_m`, SD = `r a45_sd`) consistent with the expected developmental trajectory.

All age-group means are negative relative to the Levante scale origin. This is expected given that Kimochis children (ages 2–5) fall below the Levante calibration age range of 5–12, and θ estimates are based on only 4 IRT items covering the hearts and flowers blocks only, with no mixed block items contributing to scoring.

## Comparing reliability estimates with LEVANTE
```{r}
#| label: reliability-comparison
#| include: false

# levante reliability - one row per site
levante_rxx <- task_rxx |>
  filter(task == "hf") |>
  select(site, rxx) |>
  mutate(
    sample = case_when(
      site == "pilot_mpieva_de"   ~ "Levante (Germany)",
      site == "pilot_uniandes_co" ~ "Levante (Colombia)",
      site == "pilot_western_ca"  ~ "Levante (Canada)"
    ),
    n_items = 9,
    scoring = "levante-equivalent"
  ) |>
  select(sample, n_items, rxx, scoring)

# reliability by timepoint
plus_rxx_by_time <- plus_hf_scored_collapsed |>
  group_by(time) |>
  summarise(
    n   = n(),
    rxx = round(1 - mean(se_collapsed^2) / var(theta_collapsed), 3),
    .groups = "drop"
  ) |>
  mutate(sample = "PLUS", time = as.character(time))

kimochis_rxx_by_time <- kimochis_hf_scored_collapsed |>
  group_by(time) |>
  summarise(
    n   = n(),
    rxx = round(1 - mean(se_collapsed^2) / var(theta_collapsed), 3),
    .groups = "drop"
  ) |>
  mutate(sample = "Kimochis", time = as.character(time))
```

```{r}
#| label: reliability-by-timepoint-table
#| echo: false

# build full reliability table with pooled + by-timepoint rows
rxx_full_table <- bind_rows(
  # levante rows
  levante_rxx |>
    mutate(time = NA_character_, n = NA_integer_) |>
    select(sample, time, n, rxx),

  # PLUS rows
  tibble(sample = "PLUS", time = "Pooled", n = nrow(plus_hf_scored_collapsed), rxx = plus_reliability_collapsed),
  plus_rxx_by_time |> select(sample, time, n, rxx),

  # Kimochis rows
  tibble(sample = "Kimochis", time = "Pooled", n = nrow(kimochis_hf_scored_collapsed), rxx = kimochis_reliability_collapsed),
  kimochis_rxx_by_time |> select(sample, time, n, rxx)
)

flextable(rxx_full_table) |>
  set_header_labels(sample = "Sample", time = "Timepoint", n = "N", rxx = "Reliability") |>
  colformat_double(j = "rxx", digits = 2) |>
  merge_v(j = "sample") |>
  theme_vanilla() |>
  autofit()

```

### Proportion correct by timepoint
```{r}
#| label: plus-spring-ceiling-check
#| echo: false
#| tbl-cap: "Item-level proportion correct by timepoint (PLUS)"

# item-level proportion correct by timepoint
plus_item_props <- plus_hf_collapsed |>
  filter(item_uid %in% shared_uid_types) |>
  group_by(time, item_uid) |>
  summarise(prop_correct = mean(correct_collapsed, na.rm = TRUE), .groups = "drop") |>
  mutate(time = as.character(time))

plus_item_props |>
  pivot_wider(names_from = time, values_from = prop_correct) |>
  rename(Item = item_uid) |>
  flextable() |>
  colformat_double(digits = 3) |>
  theme_vanilla() |>
  autofit()

```

```{r}
#| label: distribution-by-timepoint
#| echo: false
#| fig-cap: "Distribution of ability estimates (θ) by timepoint in PLUS, illustrating the ceiling effect in Spring"

plus_hf_scored_collapsed |>
  mutate(time = as.character(time)) |>
  ggplot(aes(x = theta_collapsed, fill = time)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Ability (θ)",
    y = "Density",
    fill = "Timepoint"
  ) +
  theme_minimal()
  
```
Reliability estimates varied substantially across timepoints. For PLUS Fall, reliability was acceptable (r = `r plus_rxx_by_time |> filter(time == "PLUS Fall") |> pull(rxx)`), but the PLUS Spring estimate was severely negative (r = `r plus_rxx_by_time |> filter(time == "PLUS Spring") |> pull(rxx)`). Negative reliability occurs when measurement error variance exceeds total score variance — that is, when scores show very little spread. This is consistent with the pronounced ceiling effect observed in PLUS Spring, where item-level proportion correct ranged from `r min(plus_item_props |> filter(time == "PLUS Spring") |> pull(prop_correct)) |> round(3)` to `r max(plus_item_props |> filter(time == "PLUS Spring") |> pull(prop_correct)) |> round(3)`. When nearly all children answer nearly all items correctly, the items no longer meaningfully differentiate between individuals, and the reliability estimate becomes unstable. The PLUS Spring scores should therefore be interpreted with caution, and the pooled reliability estimate (r = `r plus_reliability_collapsed`) is also likely deflated by this effect.

### Check item fit
```{r}
#| label: plus-fall-item-fit
#| tbl-cap: "Item-level fit statistics for PLUS Fall (S-X2 test)"
#| echo: false
#| warning: false
#| message: false

# refit model on Fall only 
plus_fall_wide <- plus_hf_collapsed_wide |>
  filter(as.character(time) == "PLUS Fall") |>
  select(all_of(shared_uid_types)) |>
  as.matrix()

# rebuild fixed parameters for fall only matrix
fixed_pars_fall <- mirt(
  data     = plus_fall_wide,
  model    = 1,
  itemtype = rep("Rasch", ncol(plus_fall_wide)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

# fit model
hf_mod_plus_fall <- mirt(
  data      = plus_fall_wide,
  model     = 1,
  itemtype  = rep("Rasch", ncol(plus_fall_wide)),
  pars      = fixed_pars_fall,
  technical = list(NCYCLES = 1)
)

# pre-compute fit stats for inline reporting and table
plus_fall_itemfit <- itemfit(hf_mod_plus_fall, fit_stats = "S_X2", na.rm = TRUE)
plus_fall_rmsea_min <- plus_fall_itemfit |> pull(RMSEA.S_X2) |> min() |> round(3)
plus_fall_rmsea_max <- plus_fall_itemfit |> pull(RMSEA.S_X2) |> max() |> round(3)

# display table
plus_fall_itemfit |>
  as_tibble() |>
  arrange(p.S_X2) |>
  mutate(p.S_X2 = format(p.S_X2, scientific = TRUE, digits = 3)) |>
  rename(Item = item, RMSEA = RMSEA.S_X2, p = p.S_X2) |>
  flextable() |>
  colformat_double(digits = 3) |>
  theme_vanilla() |>
  autofit()

```
Item fit was evaluated using the S-X2 statistic for the PLUS Fall sample (N = `r nrow(plus_fall_wide)`. All items showed significant misfit (all p < .001), with RMSEA values ranging from `r itemfit(hf_mod_plus_fall, fit_stats = "S_X2", na.rm = TRUE) |> pull(RMSEA.S_X2) |> min() |> round(3)` to `r itemfit(hf_mod_plus_fall, fit_stats = "S_X2", na.rm = TRUE) |> pull(RMSEA.S_X2) |> max() |> round(3)`. The worst-fitting item was `heartsflowers_flower_start`, which showed substantially higher misfit than the remaining items. This pattern of uniform misfit is not unexpected when item parameters are fixed from an external calibration sample rather than freely estimated: the Levante parameters were derived from children aged 5–12 assessed in a different context, and some degree of parameter drift is anticipated when these are applied to the PLUS sample. The misfit does not invalidate the scores but suggests that the fixed-parameter model fits these data less well than a freely estimated model would, and that ability estimates carry additional uncertainty beyond what the standard errors reflect.
