---
title: "Spark_h&f"
format: html
editor: visual
---
# Applying the Levante IRT Model to the PLUS Dataset
The goal of this analysis is to evaluate the Levante Hearts and Flowers (H&F) IRT model's performance on an external dataset collected by Spark Lab, as a test of the model's generalizability. The Spark Lab data consists of two studies: PLUS (3rd–5th graders, group-based tablet assessment) and Kimochis (preschoolers, 1:1 assessment). 
```{r}
#| label: packages
#| include: false
library(tidyverse)   
library(mirt)        
library(here)
library(glue)
```

```{r}
#| label: load-data
#| include: false
plus_hf_merged <- readRDS("plus_hf_merged.rds")
kimochis_hf <- readRDS("kimochis_hf.rds")
# colnames(plus_hf_merged)
# colnames(kimochis_hf)
```

```{r}
#| label: explore-data
#| include: false

head(plus_hf_merged, 3)
head(kimochis_hf, 3)

unique(plus_hf_merged$block)
unique(plus_hf_merged$stim_shape)

unique(kimochis_hf$block)
unique(kimochis_hf$stim_shape)

# check structure
glimpse(plus_hf_merged)
glimpse(kimochis_hf)

# ccount of what blocks are present
plus_hf_merged |> count(block) 
kimochis_hf |> count(block)

# check shape of PLUS
plus_hf_merged |> 
  filter(block == "flowers test") |>
  count(stim_shape, resp_side, stim_side) |>
  head(20)
```

```{r}
#| label: levante-scores
#| include: false

# install remotes
# install.packages("remotes")
# install rlevante 
# remotes::install_github("levante-framework/rlevante")

best_coefs <- read_rds(here("02_scoring_outputs/best_coefs.rds"))
mods_best  <- read_rds(here("02_scoring_outputs/mods_best.rds"))

# inspect
glimpse(best_coefs)
glimpse(mods_best)
best_coefs |> distinct(task)

# get hf item parameters 
hf_coefs <- best_coefs |> filter(task == "hf")

# see all item names
hf_coefs |> distinct(item)

# pull the hf model record
hf_mod_rec <- mods_best |> filter(task == "hf") |> pull(mod_rec) |> pluck(1)

# inspect the model record slots
isVirtualClass("ModelRecord")
slotNames(hf_mod_rec)

# look at items slot specifically
hf_mod_rec@items
```

```{r}
#| label: levante-hf-structure
#| include: false

task_rxx <- read_rds(here("02_scoring_outputs/task_rxx.rds"))
colnames(task_rxx)
task_rxx |> filter(task == "hf")

# how many items in the levante hf model per site?
best_coefs |> 
  filter(task == "hf") |> 
  group_by(group) |> 
  distinct(item) |> 
  count()

# Load data
task_data_nested <- read_rds(here("01_fetched_data/task_data_nested.rds"))
glimpse(task_data_nested$data)

# look at the data structure for hf in levante
task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  group_by(site, user_id, run_id) |>
  summarise(n_trials = n(), .groups = "drop") |>
  group_by(site) |>
  summarise(mean_trials   = mean(n_trials),
            sd_trials     = sd(n_trials),
            median_trials = median(n_trials))

task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  group_by(site, user_id, run_id) |>
  summarise(n_unique_items = n_distinct(item_uid), .groups = "drop") |>
  group_by(site) |>
  summarise(mean_unique = mean(n_unique_items),
            median_unique = median(n_unique_items))
```

```{r}
#| label: prep-plus-hf-items
#| include: false

# compute accuracy and filter PLUS to test blocks only
plus_hf_prepped <- plus_hf_merged |>
  filter(block %in% c("hearts test", "flowers test", "mixed test")) |>
  mutate(
    # accuracy rule depends on stimulus type
    correct = case_when(
      stim_shape == "heart"  ~ resp_side == stim_side, # congruent
      stim_shape == "flower" ~ resp_side != stim_side, # incongruent
      TRUE ~ NA
    ),
    # treat timeouts (NA resp_side) as incorrect
    correct = if_else(is.na(resp_side), FALSE, correct),
    correct = as.integer(correct),
    # recode block to match levante naming
    block_clean = case_when(
      block == "hearts test"  ~ "hearts",
      block == "flowers test" ~ "flowers",
      block == "mixed test"   ~ "heartsflowers"
    )
  )

# derive trial type (start / stay / switch) within each person x block
plus_hf_items <- plus_hf_prepped |>
  arrange(stud_id, time, block_clean, trial_num) |>
  group_by(stud_id, time, block_clean) |>
  mutate(
    prev_shape = lag(stim_shape),
    trial_type = case_when(
      trial_num == min(trial_num) ~ "start",
      stim_shape == prev_shape   ~ "stay",
      stim_shape != prev_shape   ~ "switch"
    )
  ) |>
  # number trials sequentially within each type
  group_by(stud_id, time, block_clean, stim_shape, trial_type) |>
  mutate(trial_type_num = row_number()) |>
  ungroup() |>
  # construct item name matching levante format
  mutate(
    item = glue("{block_clean}_{stim_shape}_{trial_type}-{trial_type_num}")
  )

# check if names match with levante?
plus_hf_items |> distinct(item) |> arrange(item)

# check overlap with levante items
levante_items <- hf_mod_rec@items
plus_items <- plus_hf_items |> distinct(item) |> pull(item)

# items in both
intersect(plus_items, levante_items)

# items in PLUS but not levante
setdiff(plus_items, levante_items)

# items in levante but not in PLUS (these will be NA in scoring)
setdiff(levante_items, plus_items)
# 57 out of 74 Levante items are present in PLUS. 

mean_timeout_pct <- plus_hf_items |>
  filter(block_clean != "hearts") |>
  group_by(stud_id, time) |>
  summarise(prop_timeout = mean(is.na(resp_side)), .groups = "drop") |>
  summarise(mean_timeout = mean(prop_timeout)) |>
  mutate(mean_timeout = round(mean_timeout * 100, 1)) |>
  pull(mean_timeout)

# flag and remove kids failing hearts attention check
plus_hearts_check <- plus_hf_items |>
  filter(block_clean == "hearts") |>
  group_by(stud_id, time) |>
  summarise(hearts_acc = mean(correct, na.rm = TRUE), .groups = "drop") |>
  filter(hearts_acc >= 0.5)

# pivot to wide format (one row per person, one col per item)
plus_hf_wide <- plus_hf_items |>
  semi_join(plus_hearts_check, by = c("stud_id", "time")) |>
  select(stud_id, time, item, correct) |>
  pivot_wider(names_from = item, values_from = correct)

```

```{r}
#| label: plus-hf-levante-equivalent
#| include: false

# collapse trial positions: strip position number from item label
# and take proportion correct per person x item_uid type
plus_hf_collapsed <- plus_hf_items |>
  semi_join(plus_hearts_check, by = c("stud_id", "time")) |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  group_by(stud_id, time, item_uid) |>
  summarise(prop_correct = mean(correct, na.rm = TRUE), .groups = "drop") |>
  mutate(correct_collapsed = as.integer(prop_correct >= 0.5))

# define levante item_uid types
levante_uid_types <- c(
  "hearts_heart_start", "hearts_heart_stay",
  "flowers_flower_start", "flowers_flower_stay",
  "heartsflowers_heart_start", "heartsflowers_flower_start",
  "heartsflowers_heart_stay", "heartsflowers_flower_stay",
  "heartsflowers_heart_switch", "heartsflowers_flower_switch"
)

# shared item_uid types between PLUS and levante
plus_uid_types <- plus_hf_collapsed |> distinct(item_uid) |> pull(item_uid)
shared_uid_types <- intersect(plus_uid_types, levante_uid_types)

# pivot to wide
plus_hf_collapsed_wide <- plus_hf_collapsed |>
  select(stud_id, time, item_uid, correct_collapsed) |>
  pivot_wider(names_from = item_uid, values_from = correct_collapsed)

# response matrix
plus_collapsed_matrix <- plus_hf_collapsed_wide |>
  select(all_of(shared_uid_types)) |>
  as.matrix()
```

```{r}
#| label: plus-hf-levante-equivalent-rasch
#| include: false

# verify g and u values first
hf_mod_rec@model_vals |>
  filter(group == "pilot_mpieva_de", name %in% c("g", "u")) |>
  group_by(name) |>
  summarise(mean = mean(value), min = min(value), max = max(value))

# levante lookup for collapsed items - d parameter only
levante_lookup_collapsed <- hf_mod_rec@model_vals |>
  filter(group == "pilot_mpieva_de",
         name == "d") |>  # Rasch only needs d
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  filter(item_uid %in% levante_uid_types) |>
  group_by(item_uid, name) |>
  slice_min(item, n = 1) |>  # use position-1 difficulty as representative
  ungroup() |>
  select(item = item_uid, name, value)

# fix parameters - Rasch
fixed_pars_collapsed <- mirt(
  data     = plus_collapsed_matrix,
  model    = 1,
  itemtype = rep("Rasch", ncol(plus_collapsed_matrix)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

# fit model
hf_mod_plus_collapsed <- mirt(
  data      = plus_collapsed_matrix,
  model     = 1,
  itemtype  = rep("Rasch", ncol(plus_collapsed_matrix)),
  pars      = fixed_pars_collapsed,
  technical = list(NCYCLES = 1)
)

# score
plus_scores_collapsed <- fscores(hf_mod_plus_collapsed, method = "EAP", full.scores.SE = TRUE)
plus_hf_scored_collapsed <- plus_hf_collapsed_wide |>
  select(stud_id, time) |>
  mutate(
    theta_collapsed = plus_scores_collapsed[, "F1"],
    se_collapsed    = plus_scores_collapsed[, "SE_F1"]
  )

plus_reliability_collapsed <- round(
  1 - mean(plus_hf_scored_collapsed$se_collapsed^2) / var(plus_hf_scored_collapsed$theta_collapsed),
  3
)
plus_reliability_collapsed

# grade summary using collapsed theta
grade_summary <- plus_hf_scored_collapsed |>
  left_join(plus_hf_merged |> distinct(stud_id, pl_grade), by = "stud_id") |>
  group_by(pl_grade) |>
  summarise(m  = round(mean(theta_collapsed, na.rm = TRUE), 2),
            sd = round(sd(theta_collapsed, na.rm = TRUE), 2),
            .groups = "drop")

g3_m  <- grade_summary |> filter(pl_grade == 3) |> pull(m)
g3_sd <- grade_summary |> filter(pl_grade == 3) |> pull(sd)
g4_m  <- grade_summary |> filter(pl_grade == 4) |> pull(m)
g4_sd <- grade_summary |> filter(pl_grade == 4) |> pull(sd)
g5_m  <- grade_summary |> filter(pl_grade == 5) |> pull(m)
g5_sd <- grade_summary |> filter(pl_grade == 5) |> pull(sd)
```

## Data preparation and scoring - PLUS 
Trial-level data from PLUS were filtered to test blocks only (hearts test, flowers test, mixed test), excluding practice trials. Accuracy was derived from the stimulus-response pairing: trials with a heart stimulus were scored correct if the response matched the stimulus side (congruent rule), and trials with a flower stimulus were scored correct if the response was on the opposite side (incongruent rule). Trials with no response (NA) were coded as incorrect. Children were excluded if they scored below 50% on the hearts block, treating this as an attention check consistent with standard Levante scoring practice. This yielded a final analytic sample of 1,430 children across fall and spring timepoints.

Each trial was assigned an item label following the Levante naming convention: `{block}_{stimulus}_{trial_type}-{trial_number}`, where trial type was classified as start (first trial in block), stay (same stimulus type as previous trial), or switch (different stimulus type from previous trial). Repeated trials of the same 
block-stimulus-trial_type combination were aggregated into a single analytic unit by taking the proportion correct across all trials of that type and binarising at a 0.5 threshold. This yielded 9 IRT items in PLUS out of the 10 analytic units used in Levante scoring. The missing item type (heartsflowers_heart_start) was absent from PLUS due to differences in block structure between the two assessments.

Ability scores (θ) showed an increasing pattern across grades (3rd grade: M = `r g3_m`, SD = `r g3_sd`; 4th grade: M = `r g4_m`, SD = `r g4_sd`; 5th grade: M = `r g5_m`, SD = `r g5_sd`), consistent with the expected developmental trajectory and supporting the construct validity of the scoring approach in this sample. Grade-level means are close to the Levante scale origin and slightly positive, suggesting that 3rd–5th graders in PLUS perform at a broadly similar level to the Levante sample. 
```{r}
#| label: prep-kimochis-hf-items
#| include: false

# Initial checks
# check stim_shape values (we saw these may be capitalized)
kimochis_hf |> count(stim_shape)

# check acc vs acc_to 
kimochis_hf |>
  filter(str_detect(block, "test")) |>
  summarise(
    mean_acc    = mean(acc, na.rm = TRUE),
    mean_acc_to = mean(acc_to, na.rm = TRUE),
    prop_timeout = mean(to, na.rm = TRUE)
  )

# check trial_num range per block
kimochis_hf |>
  filter(str_detect(block, "test")) |>
  group_by(block) |>
  summarise(max_trial = max(trial_num), min_trial = min(trial_num))
####################

kimochis_hf_items <- kimochis_hf |>
  filter(str_detect(block, "test")) |>
  mutate(
    # recode stim_shape to lowercase to match levante
    stim_shape  = str_to_lower(stim_shape),
    block_clean = case_when(
      block == "hearts test"  ~ "hearts",
      block == "flowers test" ~ "flowers"
    )
  ) |>
  arrange(id, time, block_clean, trial_num) |>
  group_by(id, time, block_clean) |>
  mutate(
    # within each block all stimuli are the same shape, so no switches
    trial_type     = if_else(trial_num == min(trial_num), "start", "stay"),
    trial_type_num = if_else(trial_type == "start", 1L, row_number() - 1L)
  ) |>
  ungroup() |>
  mutate(
    item = paste0(block_clean, "_", stim_shape, "_", trial_type, "-", trial_type_num)
  )

# check item overlap with levante
kimochis_items  <- kimochis_hf_items |> distinct(item) |> pull(item)
intersect(kimochis_items, levante_items)
setdiff(kimochis_items, levante_items)   # kimochis items not in levante
setdiff(levante_items, kimochis_items)   # levante items not in kimochis

# attention check - exclude kids scoring <50% on hearts
kimochis_hearts_check <- kimochis_hf_items |>
  filter(block_clean == "hearts") |>
  group_by(id, time) |>
  summarise(hearts_acc = mean(acc, na.rm = TRUE), .groups = "drop") |>
  filter(hearts_acc >= 0.5)
```

```{r}
#| label: kimochis-hf-levante-equivalent
#| include: false

# collapse trial positions
kimochis_hf_collapsed <- kimochis_hf_items |>
  semi_join(kimochis_hearts_check, by = c("id", "time")) |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  group_by(id, time, item_uid) |>
  summarise(prop_correct = mean(acc, na.rm = TRUE), .groups = "drop") |>
  mutate(correct_collapsed = as.integer(prop_correct >= 0.5))

# shared item_uid types between Kimochis and levante
kimochis_uid_types <- kimochis_hf_collapsed |> distinct(item_uid) |> pull(item_uid)
shared_uid_kimochis <- intersect(kimochis_uid_types, levante_uid_types)

# pivot to wide
kimochis_hf_collapsed_wide <- kimochis_hf_collapsed |>
  select(id, time, item_uid, correct_collapsed) |>
  pivot_wider(names_from = item_uid, values_from = correct_collapsed)

# response matrix
kimochis_collapsed_matrix <- kimochis_hf_collapsed_wide |>
  select(all_of(shared_uid_kimochis)) |>
  as.matrix()
```

```{r}
#| label: kimochis-hf-levante-equivalent-rasch
#| include: false

levante_lookup_collapsed_kimochis <- hf_mod_rec@model_vals |>
  filter(group == "pilot_mpieva_de", name == "d") |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  filter(item_uid %in% shared_uid_kimochis) |>
  group_by(item_uid, name) |>
  slice_min(item, n = 1) |>
  ungroup() |>
  select(item = item_uid, name, value)

fixed_pars_collapsed_kimochis <- mirt(
  data     = kimochis_collapsed_matrix,
  model    = 1,
  itemtype = rep("Rasch", ncol(kimochis_collapsed_matrix)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed_kimochis, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

hf_mod_kimochis_collapsed <- mirt(
  data      = kimochis_collapsed_matrix,
  model     = 1,
  itemtype  = rep("Rasch", ncol(kimochis_collapsed_matrix)),
  pars      = fixed_pars_collapsed_kimochis,
  technical = list(NCYCLES = 1)
)

kimochis_scores_collapsed <- fscores(hf_mod_kimochis_collapsed, method = "EAP", full.scores.SE = TRUE)
kimochis_hf_scored_collapsed <- kimochis_hf_collapsed_wide |>
  select(id, time) |>
  mutate(
    theta_collapsed = kimochis_scores_collapsed[, "F1"],
    se_collapsed    = kimochis_scores_collapsed[, "SE_F1"]
  )

kimochis_reliability_collapsed <- round(
  1 - mean(kimochis_hf_scored_collapsed$se_collapsed^2) / var(kimochis_hf_scored_collapsed$theta_collapsed),
  3
)
kimochis_reliability_collapsed

# age summary using collapsed theta
kimochis_age_summary <- kimochis_hf_scored_collapsed |>
  left_join(kimochis_hf |> distinct(id, time, age), by = c("id", "time")) |>
  mutate(age_group = cut(age, breaks = c(2, 3, 4, 5, 6),
                         labels = c("2-3", "3-4", "4-5", "5-6"))) |>
  group_by(age_group) |>
  summarise(m  = round(mean(theta_collapsed, na.rm = TRUE), 2),
            sd = round(sd(theta_collapsed, na.rm = TRUE), 2),
            .groups = "drop")

a23_m  <- kimochis_age_summary |> filter(age_group == "2-3") |> pull(m)
a23_sd <- kimochis_age_summary |> filter(age_group == "2-3") |> pull(sd)
a34_m  <- kimochis_age_summary |> filter(age_group == "3-4") |> pull(m)
a34_sd <- kimochis_age_summary |> filter(age_group == "3-4") |> pull(sd)
a45_m  <- kimochis_age_summary |> filter(age_group == "4-5") |> pull(m)
a45_sd <- kimochis_age_summary |> filter(age_group == "4-5") |> pull(sd)
```

## Data preparation and scoring - Kimochis
Trial-level data from Kimochis were filtered to test blocks only (hearts test, flowers test), excluding practice trials. Unlike PLUS, the Kimochis dataset did not include a mixed block, as the children were considered too young for the task-switching demands it requires. Accuracy was taken from the pre-computed "acc" column rather than derived from stimulus-response pairings. Timeouts were treated as missing rather than incorrect, consistent with the 1:1 assessment context in which no strict response time limit was imposed. Children were excluded if they scored below 50% on the hearts block as an attention check, consistent with standard Levante scoring practice.

Each trial was assigned an item label following the Levante naming convention. Of the 74 items in the Levante H&F model, 27 were present in Kimochis — all from the hearts and flowers blocks. The 8 additional flowers trials in Kimochis (stay-16 through stay-23) exceeded the length of the Levante flowers block and were excluded from scoring. All 46 mixed block items were absent by design. Repeated trials of the same block-stimulus-trial_type combination were aggregated into a single analytic unit by taking the proportion correct across all trials of that type and binarising at a 0.5 threshold, yielding 4 IRT items 
(hearts_heart_start, hearts_heart_stay, flowers_flower_start, flowers_flower_stay).

Ability scores (θ) showed an increasing pattern across age groups (2–3 years: M = `r a23_m`, SD = `r a23_sd`; 3–4 years: M = `r a34_m`, SD = `r a34_sd`; 4–5 years: M = `r a45_m`, SD = `r a45_sd`), consistent with the expected developmental trajectory.

All age-group means are negative relative to the Levante scale origin. This is expected given that Kimochis children (ages 2–5) fall below the Levante calibration age range of 5–12, and θ estimates are based on only 4 IRT items covering the hearts and flowers blocks only, with no mixed block items contributing to scoring.
```{r}
#| label: reliability-comparison
#| include: false

# levante reliability - one row per site
levante_rxx <- task_rxx |>
  filter(task == "hf") |>
  select(site, rxx) |>
  mutate(
    sample  = case_when(
      site == "pilot_mpieva_de"   ~ "Levante (Germany)",
      site == "pilot_uniandes_co" ~ "Levante (Colombia)",
      site == "pilot_western_ca"  ~ "Levante (Canada)"
    ),
    n_items = 9,
    scoring = "levante-equivalent"
  ) |>
  select(sample, n_items, rxx, scoring)

# spark - levante-equivalent collapsed scoring
spark_rxx_collapsed <- tibble(
  sample  = c("PLUS", "Kimochis"),
  n_items = c(length(shared_uid_types), length(shared_uid_kimochis)),
  rxx     = c(plus_reliability_collapsed, kimochis_reliability_collapsed),
  scoring = "levante-equivalent"
)

# combined table
reliability_comparison <- bind_rows(
  levante_rxx,
  spark_rxx_collapsed
) |>
  arrange(sample)

reliability_comparison
```

## Reliability
Marginal reliability was computed for each sample as 1 − mean(SE²) / var(θ), where SE is the standard error of each child's EAP θ estimate. Repeated trials of the same block-stimulus-trial_type combination were aggregated into a single 
analytic unit, yielding 9 IRT items for PLUS and 4 IRT items for Kimochis. All 768 Levante H&F runs used fixed-form administration.

Reliability for PLUS under this scoring approach was `r plus_reliability_collapsed`, which is lower than the Levante calibration estimates ranging from 0.54 (Germany) to 0.66 (Colombia). Kimochis reliability could not be meaningfully estimated under the collapsed approach, as only 4 item types were available — insufficient for stable EAP estimation.

The lower reliability in PLUS relative to Levante, despite PLUS being scored on 9 of Levante's 10 IRT items, likely reflects two factors. First, the PLUS task imposed a 750ms response time limit compared to 2,500ms in Levante, resulting in approximately `r mean_timeout_pct`% of trials being coded as incorrect due to timeouts. This systematically depresses accuracy and introduces noise into the collapsed item scores. Second, the Levante item parameters were calibrated on a sample of 5–12 year old children across three international sites, and may not transfer optimally to the PLUS population of 8-11 year-olds (3rd–5th graders) assessed under different administration conditions.

These findings suggest that the Levante H&F IRT model can be applied to external fixed-format datasets using the same collapsed scoring approach, but that reliability is sensitive to administration differences — particularly response time constraints — that affect the quality of individual item responses.