---
title: "Applying the Levante IRT Model to the PLUS and Kimochis Datasets"
author: "Fionnuala O'Reilly"
date: "`r lubridate::today()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    number-sections: true
    code-folding: hide
    theme: readable
    fig-cap-location: top
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

## Introduction
The Levante Hearts and Flowers (H&F) task is a computerized measure of inhibitory control and cognitive flexibility based on the Hearts and Flowers paradigm (Davidson et al., 2006), modelled after the AMES tablet version which has been validated across diverse global settings (Ahmed et al., 2022; Khan et al., 2024). Participants respond according to stimulus type: pressing a key on the same side as a heart (congruent rule) and on the opposite side for a flower (incongruent rule). The task includes three blocks — congruent (hearts only), incongruent (flowers only), and mixed (hearts and flowers) — with the congruent block serving as a baseline with minimal executive demands. Participants are encouraged to respond as quickly as possible; no hard time limit is imposed per trial, though responses slower than 3000ms are scored as incorrect. The task is scored using a Rasch IRT model calibrated on a cross-national sample of children aged 5–12 years.

The goal of this analysis is to evaluate the generalizability of the Levante H&F IRT model by applying it to two external datasets collected by Spark Lab. The PLUS dataset was collected from students in grades 3–5 (approximately 8–11 years) in a large urban school district in the United States, who completed the task on individual digital devices during the school day in a classroom setting. The Kimochis dataset was collected from preschool-aged children (2–5 years) in a one-to-one assessment context.

The primary generalizability test uses the PLUS dataset, which falls within the Levante calibration age range. We address this through three research questions:

1. How comparable are item difficulties across datasets? We examine whether the Levante items rank in the same order of difficulty in PLUS as in the original Levante calibration sample, using raw percent correct as a format-neutral index of item difficulty.

2. How well do Levante's item parameters transfer to PLUS? We apply fixed Levante item parameters to score PLUS children and evaluate model fit, reliability, and the plausibility of the resulting ability estimates.

3. What happens when we re-estimate item parameters from PLUS? We re-estimate item difficulty parameters freely from the PLUS Fall sample using the Levante pipeline, and compare these to the original Levante parameters to quantify how much the parameters shift and what this means for measurement quality.

We additionally examine reaction time profiles across datasets as a descriptive check on construct validity, asking whether the expected patterns of faster RTs with age and greater difficulty for incongruent trials are present in PLUS as in Levante.

Finally, we apply the model to the Kimochis dataset as a boundary condition analysis. At ages 2–5, Kimochis children fall below the Levante calibration range, and the dataset lacks a mixed block entirely. We do not expect the model to generalize to this sample, but examine how and where it breaks down as a way of characterising the limits of the model's applicability.
```{r}
#| label: packages
#| include: false
library(tidyverse)   
library(mirt)        
library(here)
library(glue)
library(flextable)
library(patchwork)
library(janitor)
library(haven)
library(dplyr)

```

```{r}
#| label: load-data
#| include: false

plus <- readRDS("plus_hf_merged.rds") |> mutate(pl_fem = as_factor(pl_fem))
kimochis <- readRDS("kimochis_hf.rds")

# run data for raw scores
run_data <- readr::read_rds(here::here("..", "levante-pilots", "01_fetched_data/run_data.rds"))
# colnames(run_data)
```

```{r}
#| label: clean-data
#| include: false

# plus cleaning
plus <- plus |>
  mutate(
    dataset    = "plus",
    age        = pl_age,
    pl_fem     = as_factor(pl_fem),
    stim_shape = str_to_lower(stim_shape),
    stim_side  = str_to_lower(stim_side),
    resp_side  = str_to_lower(resp_side),
    acc        = case_when(
      stim_shape == "heart"  ~ as.numeric(resp_side == stim_side),
      stim_shape == "flower" ~ as.numeric(resp_side != stim_side),
      TRUE                   ~ NA_real_
    ),
    acc = if_else(is.na(resp_side), 0, acc)
  ) |>
  filter(
    !block %in% c("hearts practice", "flowers practice"),
    rt >= 100, # replaces rt > 0, excludes sub-100ms trials to align with levante
    !is.na(rt),
    age <= 12 # remove children older than 11
  )

# kimochis cleaning
kimochis <- kimochis |>
  mutate(
    dataset    = "kimochis",
    stim_shape = str_to_lower(stim_shape),
    stim_side  = str_to_lower(stim_side),
    resp_side  = str_to_lower(resp_side),
    rt         = rt_all * 1000
  ) |>
  filter(!block %in% c("hearts practice", "flowers practice"))

both_descriptive <- bind_rows(
  plus     |> mutate(time = as.character(time)),
  kimochis |> mutate(time = as.character(time))
)
```

```{r}
#| label: explore-data
#| include: false

# check if any extreme RTs exist in the raw data
plus |>
  summarise(
    n_under_100  = sum(rt < 100, na.rm = TRUE),
    n_over_30000 = sum(rt > 30000, na.rm = TRUE)
  )

kimochis |>
  summarise(
    n_under_100  = sum(rt < 100, na.rm = TRUE),
    n_over_30000 = sum(rt > 30000, na.rm = TRUE)
  )

head(plus, 3)
head(kimochis, 3)

unique(plus$block)
unique(plus$stim_shape)

unique(kimochis$block)
unique(kimochis$stim_shape)

# check structure
glimpse(plus)
glimpse(kimochis)

# ccount of what blocks are present
plus |> count(block) 
kimochis |> count(block)

# check shape of PLUS
plus |> 
  filter(block == "flowers test") |>
  count(stim_shape, resp_side, stim_side) |>
  head(20)
```

## Participants
```{r}
#| label: sample-descriptives-table
#| echo: false

plus_sample <- plus |>
  group_by(time) |>
  summarise(
    dataset    = "PLUS",
    n_children = n_distinct(stud_id),
    age_mean   = mean(pl_age, na.rm = TRUE),
    age_sd     = sd(pl_age,  na.rm = TRUE),
    pct_female = mean(pl_fem == "Female", na.rm = TRUE) * 100,
    .groups    = "drop"
  ) |>
  mutate(time = as.character(time))

kimo_sample <- kimochis |>
  group_by(time) |>
  summarise(
    dataset    = "Kimochis",
    n_children = n_distinct(id),
    age_mean   = mean(age, na.rm = TRUE),
    age_sd     = sd(age,  na.rm = TRUE),
    pct_female = mean(female, na.rm = TRUE) * 100,
    .groups    = "drop"
  ) |>
  mutate(time = as.character(time))

plus_retention <- plus |>
  group_by(stud_id) |>
  summarise(n_tp = n_distinct(time)) |>
  count(n_tp) |>
  mutate(dataset = "PLUS", label = glue("{n_tp} tp: n = {n}")) |>
  group_by(dataset) |>
  summarise(retention = paste(label, collapse = "\n"), .groups = "drop")

kimo_retention <- kimochis |>
  group_by(id) |>
  summarise(n_tp = n_distinct(time)) |>
  count(n_tp) |>
  mutate(dataset = "Kimochis", label = glue("{n_tp} tp: n = {n}")) |>
  group_by(dataset) |>
  summarise(retention = paste(label, collapse = "\n"), .groups = "drop")

bind_rows(plus_sample, kimo_sample) |>
  left_join(bind_rows(plus_retention, kimo_retention), by = "dataset") |>
  mutate(
    age_m_sd   = glue("{round(age_mean, 2)} ({round(age_sd, 2)})"),
    pct_female = round(pct_female, 1)
  ) |>
  select(dataset, time, n_children, age_m_sd, pct_female, retention) |>
  rename(
    Dataset      = dataset,
    Timepoint    = time,
    N            = n_children,
    `Age M (SD)` = age_m_sd,
    `% Female`   = pct_female,
    Retention    = retention
  ) |>
  flextable() |>
  colformat_double(digits = 1) |>
  merge_v(j = c("Dataset", "Retention")) |>
  valign(j = c("Dataset", "Retention"), valign = "top") |>
  set_table_properties(width = 1, layout = "autofit") |>
  width(j = "Retention", width = 1.8) |>
  width(j = "Age M (SD)", width = 1.2) |>
  align(j = c("N", "% Female"), align = "right", part = "all") |>
  bold(part = "header") |>
  fontsize(size = 11, part = "all") |>
  theme_vanilla()
```

```{r}
#| label: kimochis-age-check
#| include: false

# are the exact same children appearing across timepoints?
kimochis |>
  group_by(id) |>
  summarise(n_timepoints = n_distinct(time)) |>
  count(n_timepoints)

# do the min/max age children appear at multiple timepoints with the same age?
kimochis |>
  filter(age == min(age, na.rm = TRUE) | age == max(age, na.rm = TRUE)) |>
  distinct(id, time, age, grade) |>
  arrange(id, time)

# is age time-varying within a child, or fixed at enrolment?
kimochis |>
  group_by(id) |>
  summarise(n_distinct_ages = n_distinct(age)) |>
  count(n_distinct_ages)

# distribution of age at each timepoint
kimochis |>
  ggplot(aes(x = age, fill = time)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(~time) +
  theme_minimal() +
  labs(title = "Age distribution by timepoint")

```

```{r}
#| label: sample-reporting-plus
#| include: false

plus_n_total <- n_distinct(plus$stud_id)
plus_n_schools <- n_distinct(plus$sch_id)

plus_summary <- plus |>
  group_by(time) |>
  summarise(
    n = n_distinct(stud_id),
    age_mean = mean(pl_age, na.rm = TRUE) |> round(2),
    age_sd   = sd(pl_age,  na.rm = TRUE)  |> round(2),
    .groups  = "drop"
  )

plus_n_fall   <- plus_summary |> filter(time == "PLUS Fall")   |> pull(n)
plus_n_spring <- plus_summary |> filter(time == "PLUS Spring") |> pull(n)
plus_age_mean <- plus_summary |> pull(age_mean) |> mean() |> round(2)
plus_age_sd   <- plus_summary |> pull(age_sd)   |> mean() |> round(2)
```
### Plus
The Plus dataset includes `r plus_n_total` unique students across two timepoints (PLUS Fall and PLUS Spring) drawn from `r plus_n_schools` schools. Sample sizes were stable across timepoints (Fall: n = `r plus_n_fall`, Spring: n = `r plus_n_spring`). Mean age was approximately `r plus_age_mean` years (SD ~`r plus_age_sd`) at both timepoints, with students in grades 3, 4, and 5 roughly evenly distributed. The sample was approximately 49% female at both timepoints.

```{r}
#| label: sample-reporting-kimochis
#| include: false

kimo_n_total <- n_distinct(kimochis$id)
kimo_n_t1 <- kimochis |> filter(time == "2021-2022 Fall")   |> summarise(n = n_distinct(id)) |> pull(n)
kimo_n_t4 <- kimochis |> filter(time == "2022-2023 Spring") |> summarise(n = n_distinct(id)) |> pull(n)

# from the sample descriptives summary table
kimo_summary <- kimochis |>
  group_by(time) |>
  summarise(
    n        = n_distinct(id),
    age_mean = mean(age, na.rm = TRUE) |> round(2),
    age_sd   = sd(age,  na.rm = TRUE)  |> round(2),
    .groups  = "drop"
  )

kimo_age_min <- kimo_summary |> summarise(m = min(age_mean)) |> pull(m)
kimo_age_max <- kimo_summary |> summarise(m = max(age_mean)) |> pull(m)

kimo_retention <- kimochis |>
  group_by(id) |>
  summarise(n_timepoints = n_distinct(time)) |>
  count(n_timepoints)

kimo_n_1tp  <- kimo_retention |> filter(n_timepoints == 1) |> pull(n)
kimo_n_2tp  <- kimo_retention |> filter(n_timepoints == 2) |> pull(n)
kimo_n_3tp  <- kimo_retention |> filter(n_timepoints == 3) |> pull(n)
kimo_n_4tp  <- kimo_retention |> filter(n_timepoints == 4) |> pull(n)

kimo_n_fixed_age <- kimochis |>
  group_by(id) |>
  summarise(n_distinct_ages = n_distinct(age)) |>
  filter(n_distinct_ages == 1) |>
  nrow()
```
### Kimochis
The Kimochis dataset includes `r kimo_n_total` unique children across four timepoints spanning two academic years (2021-2022 and 2022-2023, Fall and Spring). Sample sizes ranged from `r kimo_n_t1` children at the first timepoint to `r kimo_n_t4` at the last. The sample was approximately evenly split by sex (~46-51% female) across timepoints, with mean age ranging from `r kimo_age_min` to `r kimo_age_max` years. Children were enrolled in Pre-K and, from 2021-2022 Spring onwards, Transitional Kindergarten (TK).

Longitudinal retention was limited: the majority of children appeared at only one timepoint
(n = `r kimo_n_1tp`), with `r kimo_n_2tp` appearing at two, `r kimo_n_3tp` at three, and 
`r kimo_n_4tp` at all four timepoints. 

For the majority of children (n = `r kimo_n_fixed_age`), age did not vary across timepoints, suggesting it reflects 
approximate age at enrolment rather than age at time of assessment. Age values also appear to be rounded to the nearest month rather than recorded as a continuous measure. 

## Data processing
```{r}
#| label: data-quality-checks
#| include: false

# plus RT distribution checks
plus |>
  summarise(
    p95         = quantile(rt, 0.95),
    p99         = quantile(rt, 0.99),
    n_over_3000 = sum(rt > 3000),
    n_over_5000 = sum(rt > 5000)
  )

# kimochis: extreme RTs
kimochis |>
  filter(rt > 10000) |>
  select(id, block, trial_num, rt, acc, to) |>
  arrange(desc(rt))

# kimochis: timeout flag alignment 
# check whether to, rt NA, and acc NA are consistent
kimochis |>
  mutate(rt_na     = is.na(rt),
         acc_na    = is.na(acc),
         timed_out = to == 1) |>
  count(rt_na, acc_na, timed_out)

# plus: NA resp_side breakdown 
plus |>
  filter(is.na(resp_side)) |>
  count(block)

# kimochis: ft_stim NAs by block
kimochis |>
  filter(is.na(ft_stim)) |>
  count(block)

# kimochis: ar NAs vs acc NAs
kimochis |>
  mutate(ar_na  = is.na(ar),
         acc_na = is.na(acc)) |>
  count(ar_na, acc_na)

# missing values per column 
map_df(
  list(plus = plus, kimochis = kimochis),
  ~summarise(.x, across(everything(), ~sum(is.na(.)))),
  .id = "dataset"
)

# cell balance 
both_descriptive |>
  count(dataset, block, stim_shape, stim_side) |>
  group_by(dataset, block) |>
  mutate(expected  = mean(n),
         deviation = n - expected) |>
  filter(abs(deviation) > 5)

# duplicate rows
both_descriptive |>
  group_by(dataset) |>
  summarise(n_total    = n(),
            n_distinct = n_distinct(across(everything())),
            n_dupes    = n_total - n_distinct)

# congruency rule violations 
# for mixed test, rule is determined by stim_shape within trial
both_descriptive |>
  mutate(
    congruent          = stim_side == resp_side,
    expected_congruent = case_when(
      stim_shape == "heart"  ~ TRUE,
      stim_shape == "flower" ~ FALSE,
      TRUE                   ~ NA
    ),
    rule_violation = !is.na(expected_congruent) & (congruent != expected_congruent)
  ) |>
  count(dataset, block, rule_violation)

```

```{r}
#| label: levante-scores
#| include: false

# install remotes
# install.packages("remotes")
# install rlevante 
# remotes::install_github("levante-framework/rlevante")

best_coefs <- read_rds(here("02_scoring_outputs/best_coefs.rds"))
mods_best  <- read_rds(here("02_scoring_outputs/mods_best.rds"))

# inspect
glimpse(best_coefs)
glimpse(mods_best)
best_coefs |> distinct(task)

# get hf item parameters 
hf_coefs <- best_coefs |> filter(task == "hf")

# see all item names
hf_coefs |> distinct(item)

# pull the hf model record
hf_mod_rec <- mods_best |> filter(task == "hf") |> pull(mod_rec) |> pluck(1)

# inspect the model record slots
isVirtualClass("ModelRecord")
slotNames(hf_mod_rec)

# look at items slot specifically
hf_mod_rec@items
```

```{r}
#| label: levante-hf-structure
#| include: false

task_rxx <- read_rds(here("02_scoring_outputs/task_rxx.rds"))
colnames(task_rxx)
task_rxx |> filter(task == "hf")

# how many items in the levante hf model per site?
best_coefs |> 
  filter(task == "hf") |> 
  group_by(group) |> 
  distinct(item) |> 
  count()

# Load data
task_data_nested <- read_rds(here("01_fetched_data/task_data_nested.rds"))
glimpse(task_data_nested$data)

# look at the data structure for hf in levante
task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  group_by(site, user_id, run_id) |>
  summarise(n_trials = n(), .groups = "drop") |>
  group_by(site) |>
  summarise(mean_trials   = mean(n_trials),
            sd_trials     = sd(n_trials),
            median_trials = median(n_trials))

task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  group_by(site, user_id, run_id) |>
  summarise(n_unique_items = n_distinct(item_uid), .groups = "drop") |>
  group_by(site) |>
  summarise(mean_unique = mean(n_unique_items),
            median_unique = median(n_unique_items))
```

```{r}
#| label: prep-plus-hf-items
#| include: false

plus_hf_prepped <- plus |>
  filter(block %in% c("hearts test", "flowers test", "mixed test")) |>
  mutate(
    correct = case_when(
      stim_shape == "heart"  ~ resp_side == stim_side,
      stim_shape == "flower" ~ resp_side != stim_side,
      TRUE ~ NA
    ),
    correct = if_else(is.na(resp_side), FALSE, correct),
    correct = if_else(rt > 750, FALSE, correct),   # 750ms scoring rule
    correct = as.integer(correct),
    block_clean = case_when(
      block == "hearts test"  ~ "hearts",
      block == "flowers test" ~ "flowers",
      block == "mixed test"   ~ "heartsflowers"
    )
  )

# classify trials as start / stay / switch within each person x block
plus_hf_items <- plus_hf_prepped |>
  arrange(stud_id, time, block_clean, trial_num) |>
  group_by(stud_id, time, block_clean) |>
  mutate(
    prev_shape = lag(stim_shape),
    trial_type = case_when(
      trial_num == min(trial_num) ~ "start",
      stim_shape == prev_shape   ~ "stay",
      stim_shape != prev_shape   ~ "switch"
    )
  ) |>
  group_by(stud_id, time, block_clean, stim_shape, trial_type) |>
  mutate(trial_type_num = row_number()) |>
  ungroup() |>
  mutate(item = glue("{block_clean}_{stim_shape}_{trial_type}-{trial_type_num}"))

# timeout rate (flowers and mixed blocks only)
mean_timeout_pct <- plus_hf_items |>
  filter(block_clean != "hearts") |>
  group_by(stud_id, time) |>
  summarise(prop_timeout = mean(is.na(resp_side)), .groups = "drop") |>
  summarise(mean_timeout = round(mean(prop_timeout) * 100, 1)) |>
  pull(mean_timeout)

# attention check: exclude children scoring below 50% on hearts block
plus_hearts_check <- plus_hf_items |>
  filter(block_clean == "hearts") |>
  group_by(stud_id, time) |>
  summarise(hearts_acc = mean(correct, na.rm = TRUE), .groups = "drop") |>
  filter(hearts_acc >= 0.5)

n_hf_start_trials <- plus_hf_items |> 
  filter(block_clean == "heartsflowers", trial_type == "start") |> 
  nrow()
```

```{r}
#| label: plus-hf-levante-equivalent
#| include: false

# collapse repeated trials of same type into single binary item
plus_hf_collapsed <- plus_hf_items |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  group_by(stud_id, time, item_uid) |>
  summarise(prop_correct = mean(correct, na.rm = TRUE), .groups = "drop") |>
  mutate(correct_collapsed = as.integer(prop_correct >= 0.5))

# levante item_uid types
levante_uid_types <- c(
  "hearts_heart_start", "hearts_heart_stay",
  "flowers_flower_start", "flowers_flower_stay",
  "heartsflowers_heart_start", "heartsflowers_flower_start",
  "heartsflowers_heart_stay", "heartsflowers_flower_stay",
  "heartsflowers_heart_switch", "heartsflowers_flower_switch"
)

shared_uid_types <- intersect(
  plus_hf_collapsed |> distinct(item_uid) |> pull(item_uid),
  levante_uid_types
) |>
  setdiff("heartsflowers_heart_start")

# pivot to wide and build response matrix
plus_hf_collapsed_wide <- plus_hf_collapsed |>
  select(stud_id, time, item_uid, correct_collapsed) |>
  pivot_wider(names_from = item_uid, values_from = correct_collapsed)

plus_collapsed_matrix <- plus_hf_collapsed_wide |>
  select(all_of(shared_uid_types)) |>
  as.matrix()
```

```{r}
#| label: plus-hf-levante-equivalent-rasch
#| include: false

# pull levante d parameters (identical across sites under scalar invariance)
levante_lookup_collapsed <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  filter(item_uid %in% levante_uid_types) |>
  group_by(item_uid, name) |>
  slice_min(item, n = 1) |>
  ungroup() |>
  distinct(item = item_uid, name, value)

# fix item parameters to levante values
fixed_pars_collapsed <- mirt(
  data     = plus_collapsed_matrix,
  model    = 1,
  itemtype = rep("Rasch", ncol(plus_collapsed_matrix)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

# fit model with fixed parameters and estimate EAP scores
hf_mod_plus_collapsed <- mirt(
  data      = plus_collapsed_matrix,
  model     = 1,
  itemtype  = rep("Rasch", ncol(plus_collapsed_matrix)),
  pars      = fixed_pars_collapsed,
  technical = list(NCYCLES = 1)
)

plus_scores_collapsed <- fscores(hf_mod_plus_collapsed, method = "EAP", full.scores.SE = TRUE)

plus_hf_scored_collapsed <- plus_hf_collapsed_wide |>
  select(stud_id, time) |>
  mutate(
    theta_collapsed = plus_scores_collapsed[, "F1"],
    se_collapsed    = plus_scores_collapsed[, "SE_F1"]
  )

plus_reliability_collapsed <- round(
  1 - mean(plus_hf_scored_collapsed$se_collapsed^2) / var(plus_hf_scored_collapsed$theta_collapsed),
  3
)

# grade-level theta summary
grade_summary <- plus_hf_scored_collapsed |>
  left_join(plus |> distinct(stud_id, pl_grade), by = "stud_id") |>
  group_by(pl_grade) |>
  summarise(
    m  = round(mean(theta_collapsed, na.rm = TRUE), 2),
    sd = round(sd(theta_collapsed, na.rm = TRUE), 2),
    .groups = "drop"
  )

g3_m  <- grade_summary |> filter(pl_grade == 3) |> pull(m)
g3_sd <- grade_summary |> filter(pl_grade == 3) |> pull(sd)
g4_m  <- grade_summary |> filter(pl_grade == 4) |> pull(m)
g4_sd <- grade_summary |> filter(pl_grade == 4) |> pull(sd)
g5_m  <- grade_summary |> filter(pl_grade == 5) |> pull(m)
g5_sd <- grade_summary |> filter(pl_grade == 5) |> pull(sd)
```

### Data preparation and scoring - PLUS 
Trial-level data from PLUS were filtered to test blocks only (hearts test, flowers test, mixed test), excluding practice trials. Trials with RT < 100ms were excluded as likely accidental key presses (n = 842). Accuracy was derived from the stimulus-response pairing: trials with a heart stimulus were scored correct if the response matched the stimulus side (congruent rule), and trials with a flower stimulus were scored correct if the response was on the opposite side (incongruent rule). Trials with no response (NA) were scored as incorrect, consistent with the Levante scoring approach. Additionally, consistent with the 750ms response window enforced during PLUS data collection, trials with RT > 750ms were scored as incorrect (19.8% of trials). This yielded a final analytic sample of `r length(unique(plus_hf_scored_collapsed$stud_id))` children (`r format(nrow(plus_hf_scored_collapsed), big.mark = ",")` observations across fall and spring timepoints).

Each trial was assigned an item label following the Levante naming convention: `{block}_{stimulus}_{trial_type}-{trial_number}`, where trial type was classified as start (first trial in block), stay (same stimulus type as previous trial), or switch (different stimulus type from previous trial). Repeated trials of the same type (e.g., multiple heart stay trials) were collapsed into a single score per trial type. A child was scored as correct on that trial type if they got more than half of those trials right, and incorrect otherwise, giving one binary score per trial type per child to enter into the IRT model. Item difficulty parameters were taken from the Levante H&F Rasch model, which was estimated under scalar invariance across sites; parameters are therefore identical across all three Levante sites. This yielded `r length(shared_uid_types)` IRT items in PLUS out of the `r length(levante_uid_types)` analytic units used in Levante scoring. The missing item type (`r paste(setdiff(levante_uid_types, shared_uid_types), collapse = ", ")`) was absent from PLUS because the mixed block always began with a flower stimulus in this assessment (n = `r n_hf_start_trials` start trials, all flower), whereas the Levante mixed block begins with a heart stimulus. As a result, no child contributed a heartsflowers_heart_start trial and this analytic unit could not be scored.

Ability scores (θ) showed an increasing pattern across grades (3rd grade: M = `r g3_m`, SD = `r g3_sd`; 4th grade: M = `r g4_m`, SD = `r g4_sd`; 5th grade: M = `r g5_m`, SD = `r g5_sd`), consistent with the expected developmental trajectory and supporting the construct validity of the scoring approach in this sample. Grade-level means are negative relative to the Levante scale origin, likely reflecting the stricter 750ms response window applied in PLUS scoring, which results in a higher proportion of trials scored as incorrect compared to Levante's 3000ms rule.

```{r}
#| label: prep-kimochis-hf-items
#| include: false

# check stim_shape values (we saw these may be capitalized)
kimochis |> count(stim_shape)

# check acc vs acc_to 
kimochis |>
  filter(str_detect(block, "test")) |>
  summarise(
    mean_acc    = mean(acc, na.rm = TRUE),
    mean_acc_to = mean(acc_to, na.rm = TRUE),
    prop_timeout = mean(to, na.rm = TRUE)
  )

# check trial_num range per block
kimochis |>
  filter(str_detect(block, "test")) |>
  group_by(block) |>
  summarise(max_trial = max(trial_num), min_trial = min(trial_num))
####################

kimochis_hf_items <- kimochis |>
  filter(str_detect(block, "test")) |>
  mutate(
    stim_shape  = str_to_lower(stim_shape),
    block_clean = case_when(
      block == "hearts test"  ~ "hearts",
      block == "flowers test" ~ "flowers"
    ),
    acc = if_else(is.na(acc), 0L, as.integer(acc))  # aligning with levante where timeouts are treated as incorrect.
  ) |>
  arrange(id, time, block_clean, trial_num) |>
  group_by(id, time, block_clean) |>
  mutate(
    # no switches within block — all stimuli are the same shape
    trial_type     = if_else(trial_num == min(trial_num), "start", "stay"),
    trial_type_num = if_else(trial_type == "start", 1L, row_number() - 1L)
  ) |>
  ungroup() |>
  mutate(item = paste0(block_clean, "_", stim_shape, "_", trial_type, "-", trial_type_num))

# exclude children scoring below 50% on hearts block
kimochis_hearts_check <- kimochis_hf_items |>
  filter(block_clean == "hearts") |>
  group_by(id, time) |>
  summarise(hearts_acc = mean(acc, na.rm = TRUE), .groups = "drop") |>
  filter(hearts_acc >= 0.5)
```


```{r}
#| label: kimochis-hf-levante-equivalent
#| include: false

# collapse repeated trials of same type into single binary item
kimochis_hf_collapsed <- kimochis_hf_items |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  group_by(id, time, item_uid) |>
  summarise(prop_correct = mean(acc, na.rm = TRUE), .groups = "drop") |>
  mutate(correct_collapsed = as.integer(prop_correct >= 0.5))

shared_uid_kimochis <- intersect(
  kimochis_hf_collapsed |> distinct(item_uid) |> pull(item_uid),
  levante_uid_types
)

# pivot to wide and build response matrix
kimochis_hf_collapsed_wide <- kimochis_hf_collapsed |>
  select(id, time, item_uid, correct_collapsed) |>
  pivot_wider(names_from = item_uid, values_from = correct_collapsed)

kimochis_collapsed_matrix <- kimochis_hf_collapsed_wide |>
  select(all_of(shared_uid_kimochis)) |>
  as.matrix()

```

```{r}
#| label: kimochis-hf-levante-equivalent-rasch
#| include: false

# fix item parameters to levante values
fixed_pars_collapsed_kimochis <- mirt(
  data     = kimochis_collapsed_matrix,
  model    = 1,
  itemtype = rep("Rasch", ncol(kimochis_collapsed_matrix)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

# fit model with fixed parameters and estimate EAP scores
hf_mod_kimochis_collapsed <- mirt(
  data      = kimochis_collapsed_matrix,
  model     = 1,
  itemtype  = rep("Rasch", ncol(kimochis_collapsed_matrix)),
  pars      = fixed_pars_collapsed_kimochis,
  technical = list(NCYCLES = 1)
)

kimochis_scores_collapsed <- fscores(hf_mod_kimochis_collapsed, method = "EAP", full.scores.SE = TRUE)

kimochis_hf_scored_collapsed <- kimochis_hf_collapsed_wide |>
  select(id, time) |>
  mutate(
    theta_collapsed = kimochis_scores_collapsed[, "F1"],
    se_collapsed    = kimochis_scores_collapsed[, "SE_F1"]
  )

kimochis_reliability_collapsed <- round(
  1 - mean(kimochis_hf_scored_collapsed$se_collapsed^2) / var(kimochis_hf_scored_collapsed$theta_collapsed),
  3
)

# age-group theta summary
kimochis_age_summary <- kimochis_hf_scored_collapsed |>
  left_join(kimochis |> distinct(id, time, age), by = c("id", "time")) |>
  mutate(age_group = cut(age, breaks = c(2, 3, 4, 5, 6),
                         labels = c("2-3", "3-4", "4-5", "5-6"))) |>
  group_by(age_group) |>
  summarise(
    m  = round(mean(theta_collapsed, na.rm = TRUE), 2),
    sd = round(sd(theta_collapsed, na.rm = TRUE), 2),
    .groups = "drop"
  )

a23_m  <- kimochis_age_summary |> filter(age_group == "2-3") |> pull(m)
a23_sd <- kimochis_age_summary |> filter(age_group == "2-3") |> pull(sd)
a34_m  <- kimochis_age_summary |> filter(age_group == "3-4") |> pull(m)
a34_sd <- kimochis_age_summary |> filter(age_group == "3-4") |> pull(sd)
a45_m  <- kimochis_age_summary |> filter(age_group == "4-5") |> pull(m)
a45_sd <- kimochis_age_summary |> filter(age_group == "4-5") |> pull(sd)
```

```{r}
#| label: item-overlap-table
#| echo: false
#| tbl-cap: "Item coverage across Levante, PLUS, and Kimochis"

# total levante items
n_levante_total <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  distinct(item) |>
  nrow()

# levante items by block
n_levante_hearts  <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  distinct(item) |>
  filter(str_detect(item, "^hearts_")) |>
  nrow()

n_levante_flowers <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  distinct(item) |>
  filter(str_detect(item, "^flowers_")) |>
  nrow()

n_levante_mixed <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  distinct(item) |>
  filter(str_detect(item, "^heartsflowers_")) |>
  nrow()

# PLUS: levante-side raw item count matching shared uid types
n_plus_raw <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  distinct(item) |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  filter(item_uid %in% shared_uid_types) |>
  nrow()

# Kimochis: levante-side raw item count matching shared uid types
n_kimo_raw <- hf_mod_rec@model_vals |>
  filter(name == "d") |>
  distinct(item) |>
  filter(!str_detect(item, "^heartsflowers_")) |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  filter(item_uid %in% shared_uid_kimochis) |>
  nrow()

# missing IRT unit types
missing_plus <- setdiff(levante_uid_types, shared_uid_types)
missing_kimo <- setdiff(levante_uid_types, shared_uid_kimochis)

tibble(
  ` ` = c(
    "Total raw items",
    "  — Hearts block",
    "  — Flowers block",
    "  — Mixed block",
    "Raw items matched to dataset",
    "Reason(s) for non-overlap",
    "IRT analytic units"
  ),
  `Levante` = c(
    n_levante_total,
    n_levante_hearts,
    n_levante_flowers,
    n_levante_mixed,
    n_levante_total,
    "—",
    length(levante_uid_types)
  ),
  `PLUS` = c(
    n_levante_total,
    n_levante_hearts,
    n_levante_flowers,
    n_levante_mixed,
    n_plus_raw,
    paste0("Missing unit type: ", paste(missing_plus, collapse = ", ")),
    length(shared_uid_types)
  ),
  `Kimochis` = c(
    n_levante_total,
    n_levante_hearts,
    n_levante_flowers,
    n_levante_mixed,
    n_kimo_raw,
    paste0("No mixed block (", n_levante_mixed, " items); ",
           "8 flowers trials exceed Levante block length (stay-16 to stay-23)"),
    length(shared_uid_kimochis)
  )
) |>
  flextable() |>
  theme_vanilla() |>
  autofit() |>
  bold(part = "header") |>
  bold(i = c(1, 7), part = "body") |>
  italic(i = c(2, 3, 4), part = "body") |>
  hline(i = c(4, 5, 6), part = "body") |>
  fontsize(size = 11, part = "all")
```

### Data preparation and scoring - Kimochis
Trial-level data from Kimochis were filtered to test blocks only (hearts test, flowers test), excluding practice trials. The Kimochis dataset did not include a mixed block, as the children were considered too young for the task-switching demands it requires. Accuracy was taken from the pre-computed "acc" column rather than derived from stimulus-response pairings. Timeout trials (14.5% of test trials) were scored as incorrect, consistent with the Levante scoring approach. This yielded a final analytic sample of `r length(unique(kimochis_hf_scored_collapsed$id))` children (`r format(nrow(kimochis_hf_scored_collapsed), big.mark = ",")` observations) across two academic year cohorts (2021–2022 and 2022–2023), each with fall and spring timepoints.

Each trial was assigned an item label following the Levante naming convention. Of the `r n_levante_total` items in the Levante H&F model, `r n_kimo_raw` were present in Kimochis — all from the hearts and flowers blocks. The 8 additional flowers trials in Kimochis (stay-16 through stay-23) exceeded the length of the Levante flowers block and were excluded from scoring. All `r n_levante_mixed` mixed block items were absent by design. Repeated trials of the same block-stimulus-trial_type combination were aggregated into a single analytic unit: a child was scored as correct on that trial type if they got more than half of those trials right, and incorrect otherwise. This yielded `r length(shared_uid_kimochis)` IRT items (`r paste(shared_uid_kimochis, collapse = ", ")`).

Ability scores (θ) showed an increasing pattern across age groups (2–3 years: M = `r a23_m`, SD = `r a23_sd`; 3–4 years: M = `r a34_m`, SD = `r a34_sd`; 4–5 years: M = `r a45_m`, SD = `r a45_sd`) consistent with the expected developmental trajectory.

All age-group means are negative relative to the Levante scale origin. This is expected given that Kimochis children (ages 2–5) fall below the Levante calibration age range of 5–12, and θ estimates are based on only 4 IRT items covering the hearts and flowers blocks only, with no mixed block items contributing to scoring.

```{r}
#| label: sample-size-table
#| include: false

sample_sizes <- plus_hf_scored_collapsed |>
  count(time) |>
  mutate(sample = "PLUS", time = as.character(time)) |>
  bind_rows(
    kimochis_hf_scored_collapsed |>
      count(time) |>
      mutate(sample = "Kimochis", time = as.character(time))
  ) |>
  select(Sample = sample, Timepoint = time, N = n)

flextable(sample_sizes) |>
  merge_v(j = "Sample") |>
  theme_vanilla() |>
  autofit()
```

## How does raw performance compare across datasets?
```{r}
#| label: age-raw-accuracy-prep
#| include: false

# levante: proportion correct per child per run
levante_acc <- task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  group_by(site, user_id, run_id) |>
  summarise(
    prop_correct = mean(correct, na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(
    run_data |> select(user_id, run_id, age) |> mutate(age = as.numeric(age)),
    by = c("user_id", "run_id")
  ) |>
  mutate(dataset = case_when(
    site == "pilot_mpieva_de"   ~ "Levante (Germany)",
    site == "pilot_uniandes_co" ~ "Levante (Colombia)",
    site == "pilot_western_ca"  ~ "Levante (Canada)"
  ))

# plus: proportion correct per child per timepoint
plus_acc <- plus_hf_items |>
  group_by(stud_id, time) |>
  summarise(
    prop_correct = mean(correct, na.rm = TRUE),
    age = mean(age, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(dataset = as.character(time))
```

```{r}
#| label: fig-age-raw-accuracy-plot
#| echo: false
#| message: false
#| fig-cap: "Age-related raw proportion correct on the Hearts and Flowers task across datasets. Each point represents one participant run; the blue line shows a GAM-smoothed trend with 95% confidence interval. PLUS data are split by timepoint (Fall and Spring). Note that PLUS proportion correct reflects the stricter 750ms response window applied during data collection."

# combine and plot
bind_rows(
  levante_acc |> select(dataset, age, prop_correct),
  plus_acc    |> select(dataset, age, prop_correct)
) |>
  filter(!is.na(age)) |>
  mutate(dataset = factor(dataset, levels = c(
    "Levante (Germany)", "Levante (Colombia)", "Levante (Canada)",
    "PLUS Fall", "PLUS Spring"
  ))) |>
  ggplot(aes(x = age, y = prop_correct)) +
  geom_point(alpha = 0.3, size = 0.8, colour = "grey40") +
  geom_smooth(method = "gam", colour = "blue", se = TRUE) +
  facet_wrap(~dataset, nrow = 1) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(
    x = "Age (years)",
    y = "Proportion correct",
    caption = "Each point = participant run; trend = GAM per dataset"
  ) +
  theme_minimal()

```


```{r}
#| label: item-difficulty-raw-scores-prep
#| include: false

# levante: raw trial-level accuracy per item_uid
levante_item_props <- task_data_nested |>
  filter(item_task == "hf") |>
  unnest(data) |>
  mutate(item_uid = str_remove(item_uid, "-\\d+$")) |>
  filter(item_uid %in% levante_uid_types) |>
  group_by(site, item_uid) |>
  summarise(
    prop_correct = mean(correct, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    dataset = case_when(
      site == "pilot_mpieva_de"   ~ "Levante (Germany)",
      site == "pilot_uniandes_co" ~ "Levante (Colombia)",
      site == "pilot_western_ca"  ~ "Levante (Canada)"
    )
  ) |>
  select(dataset, item_uid, prop_correct)

# plus: raw trial-level accuracy per item_uid type (no collapsing)
plus_item_props_raw <- plus_hf_items |>
  mutate(item_uid = str_remove(item, "-\\d+$")) |>
  filter(item_uid %in% levante_uid_types) |>
  filter(item_uid != "heartsflowers_heart_start") |>  # exclude anomalous item
  group_by(time, item_uid) |>
  summarise(
    prop_correct = mean(correct, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(dataset = case_when(
    as.character(time) == "PLUS Fall"   ~ "PLUS Fall",
    as.character(time) == "PLUS Spring" ~ "PLUS Spring"
  )) |>
  select(dataset, item_uid, prop_correct)

# combine and plot
item_props_combined <- bind_rows(levante_item_props, plus_item_props_raw)

item_order <- levante_item_props |>
  group_by(item_uid) |>
  summarise(mean_prop = mean(prop_correct), .groups = "drop") |>
  arrange(mean_prop) |>
  pull(item_uid)

```

```{r}
#| label: fig-item-difficulty-raw-scores-plot
#| echo: false
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Raw proportion correct per item type across datasets, computed from trial-level data prior to IRT scoring. Items are ordered by mean proportion correct across Levante sites (easiest at top). Note that heartsflowers_heart_start is shown for Levante only, as this item does not exist in the PLUS assessment design."

item_props_combined |>
  mutate(
    item_uid = factor(item_uid, levels = item_order),
    dataset  = factor(dataset, levels = c(
      "Levante (Germany)", "Levante (Colombia)", "Levante (Canada)",
      "PLUS Fall", "PLUS Spring"
    ))
  ) |>
  ggplot(aes(x = prop_correct, y = item_uid, colour = dataset, shape = dataset)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  scale_colour_manual(values = c(
    "Levante (Germany)"  = "#E41A1C",
    "Levante (Colombia)" = "#377EB8",
    "Levante (Canada)"   = "#4DAF4A",
    "PLUS Fall"          = "#FF7F00",
    "PLUS Spring"        = "#984EA3"
  )) +
  scale_shape_manual(values = c(
    "Levante (Germany)"  = 16,
    "Levante (Colombia)" = 17,
    "Levante (Canada)"   = 15,
    "PLUS Fall"          = 3,
    "PLUS Spring"        = 4
  )) +
  labs(
    x      = "Proportion correct",
    y      = NULL,
    colour = "Dataset",
    shape  = "Dataset",
    title  = "Item-level proportion correct across datasets"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.margin = margin(l = 20, unit = "mm")
  )

```

We first examined raw proportion correct across datasets as an index of item difficulty, prior to any IRT modelling. @fig-age-raw-accuracy-plot shows overall accuracy as a function of age for each dataset. Across all three Levante sites, accuracy increased markedly with age, rising from approximately 50–60% at age 5 to 90–100% by age 12. PLUS Fall showed a similar upward trend across the grade 3–5 age range, though at a lower overall level than comparable Levante ages, consistent with the stricter 750ms response window applied during PLUS data collection. 

@fig-item-difficulty-raw-scores-plot shows raw proportion correct broken down by item type. The item difficulty ordering was broadly consistent across datasets — hearts and flowers block items were easiest across all datasets, while mixed block items were harder, consistent with the greater executive demands of task-switching. PLUS Fall was systematically lower than all three Levante sites on mixed block items, again likely reflecting the 750ms scoring rule, which disproportionately affects slower mixed block trials. One exception was `heartsflowers_flower_start`, where PLUS Fall showed substantially lower accuracy (~35%) relative to Levante sites (~75–85%), suggesting this item may function differently in the PLUS assessment context. This item is examined further in the item fit analysis in Section 4.

## How well do Levante's item parameters transfer to PLUS?
```{r}
#| label: reliability-comparison
#| include: false

# levante reliability - one row per site
levante_rxx <- task_rxx |>
  filter(task == "hf") |>
  select(site, rxx) |>
  mutate(
    sample = case_when(
      site == "pilot_mpieva_de"   ~ "Levante (Germany)",
      site == "pilot_uniandes_co" ~ "Levante (Colombia)",
      site == "pilot_western_ca"  ~ "Levante (Canada)"
    ),
    n_items = 9,
    scoring = "levante-equivalent"
  ) |>
  select(sample, n_items, rxx, scoring)

# reliability by timepoint
plus_rxx_by_time <- plus_hf_scored_collapsed |>
  group_by(time) |>
  summarise(
    n   = n(),
    rxx = round(1 - mean(se_collapsed^2) / var(theta_collapsed), 3),
    .groups = "drop"
  ) |>
  mutate(sample = "PLUS", time = as.character(time))

```

```{r}
#| label: reliability-by-timepoint-table
#| echo: false

# build full reliability table with pooled + by-timepoint rows
rxx_full_table <- bind_rows(
  # levante rows
  levante_rxx |>
    mutate(time = NA_character_, n = NA_integer_) |>
    select(sample, time, n, rxx),
  # PLUS rows
  tibble(sample = "PLUS", time = "Pooled", n = nrow(plus_hf_scored_collapsed), rxx = plus_reliability_collapsed),
  plus_rxx_by_time |> select(sample, time, n, rxx)
)

flextable(rxx_full_table) |>
  set_header_labels(sample = "Sample", time = "Timepoint", n = "N", rxx = "Reliability") |>
  colformat_double(j = "rxx", digits = 2) |>
  merge_v(j = "sample") |>
  theme_vanilla() |>
  autofit()

```

```{r}
#| label: plus-spring-check
#| echo: false
#| tbl-cap: "Item-level proportion correct by timepoint (PLUS)"

# item-level proportion correct by timepoint
plus_item_props <- plus_hf_collapsed |>
  filter(item_uid %in% shared_uid_types) |>
  group_by(time, item_uid) |>
  summarise(prop_correct = mean(correct_collapsed, na.rm = TRUE), .groups = "drop") |>
  mutate(time = as.character(time))

plus_item_props |>
  pivot_wider(names_from = time, values_from = prop_correct) |>
  rename(Item = item_uid) |>
  flextable() |>
  colformat_double(digits = 3) |>
  theme_vanilla() |>
  autofit()

```

Reliability estimates are reported as marginal reliability, computed as 1 minus the ratio of mean error variance to total score variance. Levante reliability ranged from `r levante_rxx |> pull(rxx) |> min() |> round(2)` to `r levante_rxx |> pull(rxx) |> max() |> round(2)` across the three sites, reflecting moderate reliability. PLUS reliability was modest but positive at both timepoints (Fall: r = `r plus_rxx_by_time |> filter(time == "PLUS Fall") |> pull(rxx)`, Spring: r = `r plus_rxx_by_time |> filter(time == "PLUS Spring") |> pull(rxx)`), and notably lower than Levante.  The lower reliability relative to Levante likely reflects the stricter 750ms response window and the group-based assessment format rather than a fundamental problem with the IRT model.

```{r}
#| label: distribution-by-timepoint
#| echo: false
#| fig-cap: "Distribution of ability estimates (θ) by timepoint in PLUS"

plus_hf_scored_collapsed |>
  mutate(time = as.character(time)) |>
  ggplot(aes(x = theta_collapsed, fill = time)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Ability (θ)",
    y = "Density",
    fill = "Timepoint"
  ) +
  theme_minimal()
  
```

### Check item fit
```{r}
#| label: plus-fall-item-fit
#| tbl-cap: "Item-level fit statistics for PLUS Fall (S-X2 test)"
#| echo: false
#| warning: false
#| message: false

# refit model on Fall only 
plus_fall_wide <- plus_hf_collapsed_wide |>
  filter(as.character(time) == "PLUS Fall") |>
  select(all_of(shared_uid_types)) |>
  as.matrix()

# rebuild fixed parameters for fall only matrix
fixed_pars_fall <- mirt(
  data     = plus_fall_wide,
  model    = 1,
  itemtype = rep("Rasch", ncol(plus_fall_wide)),
  pars     = "values"
) |>
  left_join(levante_lookup_collapsed, by = c("item", "name"), suffix = c("", "_levante")) |>
  mutate(
    value = if_else(!is.na(value_levante), value_levante, value),
    est   = FALSE
  ) |>
  select(-value_levante)

# fit model
hf_mod_plus_fall <- mirt(
  data      = plus_fall_wide,
  model     = 1,
  itemtype  = rep("Rasch", ncol(plus_fall_wide)),
  pars      = fixed_pars_fall,
  technical = list(NCYCLES = 1)
)

# pre-compute fit stats for inline reporting and table
plus_fall_itemfit <- itemfit(hf_mod_plus_fall, fit_stats = "S_X2", na.rm = TRUE)
plus_fall_rmsea_min <- plus_fall_itemfit |> pull(RMSEA.S_X2) |> min() |> round(3)
plus_fall_rmsea_max <- plus_fall_itemfit |> pull(RMSEA.S_X2) |> max() |> round(3)

# display table
plus_fall_itemfit |>
  as_tibble() |>
  arrange(p.S_X2) |>
  mutate(p.S_X2 = format(p.S_X2, scientific = TRUE, digits = 3)) |>
  rename(Item = item, RMSEA = RMSEA.S_X2, p = p.S_X2) |>
  flextable() |>
  colformat_double(digits = 3) |>
  theme_vanilla() |>
  autofit()

```
Item fit was evaluated using the S-X2 statistic for the PLUS Fall sample (N = `r nrow(plus_fall_wide)`). All items showed significant misfit (all p < .001), with RMSEA values ranging from `r plus_fall_rmsea_min` to `r plus_fall_rmsea_max`. The worst-fitting item was `heartsflowers_flower_start`, which showed substantially higher misfit than the remaining items.

## What happens when we re-estimate item parameters from PLUS?
```{r}

# New chunk - to be added later
```

## How do RTs change with age across datasets?
```{r}
#| label: trial-descriptive-stats
#| include: false

# structure
glimpse(plus)
glimpse(kimochis)

# numeric summary
plus |> select(where(is.numeric)) |> summary()
kimochis |> select(where(is.numeric)) |> summary()

# trial count by block
both_descriptive |>
  count(dataset, block) |>
  pivot_wider(names_from = dataset, values_from = n)

# accuracy by block
both_descriptive |>
  group_by(dataset, block) |>
  summarise(
    n        = n(),
    acc_mean = mean(acc, na.rm = TRUE),
    acc_sd   = sd(acc,  na.rm = TRUE),
    .groups  = "drop"
  )

```

```{r}
#| label: descriptive-table
#| echo: false

both_descriptive |>
  filter(!is.na(rt)) |>
  group_by(dataset, block) |>
  summarise(
    n_trials = n(),
    age_mean = mean(age, na.rm = TRUE),
    age_sd   = sd(age,  na.rm = TRUE),
    acc_mean = mean(acc, na.rm = TRUE),
    acc_sd   = sd(acc,  na.rm = TRUE),
    rt_mean  = mean(rt),
    rt_med   = median(rt),
    rt_sd    = sd(rt),
    rt_min   = min(rt),
    rt_max   = max(rt),
    .groups  = "drop"
  ) |>
  mutate(age_m_sd = glue("{round(age_mean, 2)} ({round(age_sd, 2)})")) |>
  select(-age_mean, -age_sd) |>
  arrange(dataset, block) |>
  rename(
    Dataset     = dataset,
    Block       = block,
    `N Trials`  = n_trials,
    `Age M (SD)` = age_m_sd,
    `Acc Mean`  = acc_mean,
    `Acc SD`    = acc_sd,
    `RT Mean`   = rt_mean,
    `RT Median` = rt_med,
    `RT SD`     = rt_sd,
    `RT Min`    = rt_min,
    `RT Max`    = rt_max
  ) |>
  flextable() |>
  colformat_double(digits = 2) |>
  merge_v(j = "Dataset") |>
  theme_vanilla() |>
  autofit()

```

```{r}
#| label: plots
#| echo: false

# RT distributions by block
p_rt_dist <- both_descriptive |>
  filter(!is.na(rt)) |>
  ggplot(aes(x = rt, fill = dataset)) +
  geom_histogram(bins = 60, alpha = 0.7, position = "identity") +
  facet_wrap(~block, scales = "free_y") +
  labs(title = "RT distributions by block", x = "RT (ms)", y = "count") +
  theme_minimal()

# RT boxplots by block x stim_shape
p_rt_box <- both_descriptive |>
  filter(!is.na(rt)) |>
  ggplot(aes(x = block, y = rt, fill = stim_shape)) +
  geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
  facet_wrap(~dataset) +
  labs(title = "RT by block × stim_shape", y = "RT (ms)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

# accuracy by block x stim_shape
p_acc <- both_descriptive |>
  group_by(dataset, block, stim_shape) |>
  summarise(acc = mean(acc, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = block, y = acc, fill = stim_shape)) +
  geom_col(position = "dodge") +
  facet_wrap(~dataset) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Accuracy by block × stim_shape", y = "proportion correct") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

print(p_rt_dist)
print(p_rt_box)
print(p_acc)

```

## Boundary condition — Kimochis
```{r}
kimochis_rxx_by_time <- kimochis_hf_scored_collapsed |>
  group_by(time) |>
  summarise(
    n   = n(),
    rxx = round(1 - mean(se_collapsed^2) / var(theta_collapsed), 3),
    .groups = "drop"
  ) |>
  mutate(sample = "Kimochis", time = as.character(time))


```
<!--
Kimochis reliability estimates were consistently negative across all timepoints (range: `r kimochis_rxx_by_time |> pull(rxx) |> min()` to `r kimochis_rxx_by_time |> pull(rxx) |> max()`). These values should not be interpreted as evidence of poor measurement quality per se, but rather reflect two fundamental limitations of applying this scoring approach to the Kimochis sample. First, scoring is based on only 4 IRT items — far fewer than the 9–10 used in PLUS and Levante — providing very limited information per child. Second, the Levante item parameters were calibrated on children aged 5–12, whereas Kimochis children are aged 2–5. The fixed difficulty parameters are therefore misaligned with the ability range of this sample, meaning the items do not function as intended for this age group.
-->










