---
execute: 
  eval: false
---

Data pre-processing happens in Redivis workflows in the LEVANTE pilots study (https://redivis.com/workspace/studies/2045/workflows).

Eventually there should be API endpoints to rerun the workflows programmatically that can be called here, for now the workflows need to be run manually.

Their code is reproduced here for clarity/debugging purpose. It's identical across all pilot datasets except that for `us_downward_extension_pilot`, `remove_incomplete_runs` is `FALSE`.

pre-install:

```{bash}
R -e 'devtools::install_github("levante-framework/rlevante")'
```

trial data:

```{r}
trials <- rlevante::get_trials(rlevante::notebook_dataset(redivis),
                               remove_incomplete_runs = TRUE,
                               remove_invalid_runs = TRUE,
                               remove_invalid_trials = FALSE)

redivis$current_notebook()$create_output_table(trials, name = "trial_data_output")
```


survey data:

```{r}
surveys <- rlevante::get_surveys(rlevante::notebook_dataset(redivis))

redivis$current_notebook()$create_output_table(surveys, name = "survey_data_output")
```
