---
title: "LEVANTE EF Tasks - Interim Report"
author: "Mike Frank"
toc: true
number-sections: true
highlight-style: pygments
format:
  html: 
    code-fold: true
---


Required packages

```{r}

#| echo: false

library("tidyverse")
library("glue")
library("here")
library("viridis")
library("purrr")
library("readr")
library("dplyr")
library("tidyr")
library("quarto")
library("forcats")
library("ggplot2")

```

Helper functions

```{r}
#| echo: false

source(here("02_score_data","irt_helpers.R"))
source(here::here("plot_settings.R"))
source(here("03_explore_tasks/explore_helper.R"))

```

Get EF data

```{r}
ef <- load_task_data(c("hearts-and-flowers","same-different-selection","memory-game"))
ef

```

Load multigroup models

```{r}
best_multigroup <- readRDS(here("02_scoring_outputs", "irt_outputs", "item_coefs.rds"))
multigroup_scores <- readRDS(here("02_scoring_outputs", "scores", "scores_multigroup.rds"))
```


First plot sumscores

```{r}
ef_runs <- ef |>
  filter(site != "us_pilot") |>
  group_by(site, user_id, run_id, task_id) |>
  summarise(
    correct = mean(correct),
    age = mean(age),
    .groups = "drop"
  )

ef_sum_scores <- ggplot(ef_runs, aes(x = age, y = correct)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "loess", lambda = 1) +
  facet_grid(site ~ task_id) + 
  ylim(0,1) +
  labs(
    title = "Age-related accuracy, smoothed trends by site",
    x = "Age (years)",
    y = "Proportion correct",
    caption = "Note: Each point represents a participant run; smoothed trend by site"
  )

print(ef_sum_scores)

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "ef_sum_scores.png"),
  plot = ef_sum_scores,
  width = 10,
  height = 6,
  dpi = 300
)
```

Now go to task analysis. 

# Hearts and Flowers

Zoom in on HF. 

```{r}
hf <- ef |>
  filter(task_id == "hearts-and-flowers") 
```

Look at all phases. 

## Sumscore

```{r}
hf_blocks <- hf |>
  filter(site != "us_pilot") |>
  group_by(site, user_id, run_id, task_id, item_group) |>
  summarise(correct = mean(correct), 
            age = mean(age), 
            n = n())

ggplot(hf_blocks,
       aes(x = age, y = correct)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  geom_hline(yintercept = .5, lty = 2) + 
  ylim(0,1) + 
  facet_grid(site ~ item_group)
```

F notes that there are a lot of younger kids BELOW chance. Let's see if they are missing data. 

```{r}
ggplot(hf_blocks, aes(x = n, y = correct)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_grid(site ~ item_group)
```

NO - this is not because of missing data. Something else is happening. 

```{r}
hf_runs <- hf |>
  filter(site != "us_pilot") |>
  group_by(site, user_id, run_id, task_id) |>
  summarise(correct_rt = mean(as.numeric(rt[correct]), na.rm=TRUE),
            correct = mean(correct),
            age = mean(age))

```

Histograms. 

```{r}
ggplot(filter(ef_runs, task_id == "hearts-and-flowers") |>
         mutate(age_group = cut(age, c(5, 7, 9, 11, 13, include.lowest=TRUE))) |>
         filter(!is.na(age_group)),
       aes(x = correct)) + 
  geom_histogram(binwidth = .25) +
  geom_vline(xintercept = .5, lty = 2) +
  scale_x_continuous(breaks = c(0,.25, .5, .75, 1)) + 
  facet_grid(site ~ age_group)
```

Reaction time. 

```{r}
ggplot(hf_runs,
       aes(x = age, y = correct_rt)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, lty = 2) + 
  # ylim(0,1) + 
  facet_grid(site ~ task_id)
```

```{r}
ggplot(hf, aes(x = as.numeric(rt))) + 
  geom_histogram() + 
  scale_x_log10(breaks = c(100,200, 500,1000,2000,3000,5000,10000)) + coord_flip()
```


## IRT

We now use the multigroup IRT models. 

```{r}

run_ages <- ef |>
  filter(site != "us_pilot") |>
  select(site, run_id, user_id, age) |>
  distinct()

multigroup_scores_hf <- multigroup_scores |>
  filter(task_id == "hearts-and-flowers", site != "us_pilot") |>
  select(site, task_id, user_id, run_id, metric_type, metric_value ) |>
  left_join(run_ages)

ef_irt <- ggplot(multigroup_scores_hf, aes(x = age, y = metric_value, col = site)) + 
  geom_point() + 
  geom_smooth() +
  labs(
    title = "Developmental trends in EF ability by site",
    x = "Age (years)",
    y = "Ability (IRT Score)",
    colour = "Site"
  )
print(ef_irt)

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "ef_irt.png"),
  plot = ef_irt,
  width = 10,
  height = 6,
  dpi = 300
)
```

Look at coefficients. 

Hearts and flowers - IRT scores by trial type

```{r}

# Hearts and flowers - IRT scores by trial type
hf_coefs <- best_multigroup |>
  filter(task_id == "hearts-and-flowers", term == "difficulty", site != "us_pilot") |>
  separate(item, into = c("task", "block", "trial", "number"), sep = "_", remove = FALSE) |>
  select(-number)

# Plot
library(ggrepel)
hf_trial_type <- ggplot(hf_coefs, aes(x = block, y = value, colour = trial, shape = trial)) +
  geom_point(size = 3) +
  geom_text_repel(aes(label = trial), size = 3, show.legend = FALSE) +
  ylab("IRT difficulty") +
  xlab("Block") +
  facet_wrap(~site) +
  labs(title = "Hearts and Flowers: Item difficulty by block and trial type") 

print(hf_trial_type)

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "hf_trial_type.png"),
  plot = hf_trial_type,
  width = 10,
  height = 6,
  dpi = 300
)



# Mike previous code
hf_coefs <- filter(best_multigroup, task_id == "hearts-and-flowers")$coefs[[1]] |>
  separate(item, sep = "_", into = c("task","block","trial", "number")) |>
  select(-number) |>
  distinct()

ggplot(hf_coefs, aes(x = a1, y = -d, col = block, pch = trial)) + 
  geom_point() + 
  geom_text_repel(aes(label = trial)) + 
  xlab("Slope") + 
  ylab("Difficulty") + 
  facet_wrap(~site) 


```

# Memory game

```{r}
mg <- ef |>
  filter(task_id == "memory-game", site != "us_pilot") |>
  separate(item_uid , into = c("task", "direction", "size", "span"), sep = "_") |>
  mutate(span = as.numeric(str_replace(span, "len",""))) |>
  group_by(run_id, direction) |>
  mutate(max_span = max(span))
```

## Sumscore

```{r}
mg_spans <- mg |>
  filter(site != "us_pilot") |>
  group_by(site, span, size, direction) |>
  summarise(n = n(), 
            correct = mean(correct, na.rm = TRUE))

mg_span_raw <- ggplot(mg_spans, aes(x = span, y = correct, col = direction, pch = size)) + 
  geom_point(aes(size = n)) + 
  geom_line(aes(lty = size)) + 
  facet_wrap(~site) +
  labs(
  title = "Memory Game: Raw accuracy by span length, direction, and grid size",
  x = "Span length",
  y = "Proportion correct"
  )

print(mg_span_raw)

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "mg_span_raw.png"),
  plot = mg_span_raw,
  width = 10,
  height = 6,
  dpi = 300
)

```

```{r}
mg |>
  filter(direction == "forward", span > 1, site != "us_pilot") |>
  mutate(age_rounded = as_factor(round(age))) |>
  group_by(age_rounded, span) |>
  summarise(correct = mean(correct), 
            n = n()) |>
  ggplot(aes(x = span, y = correct, col = age_rounded)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(method = "lm", se =FALSE)

```

```{r}
mg |>
  filter(direction == "forward", span > 1, site != "us_pilot") |>
  mutate(max_span = as_factor(max_span)) |>
  group_by(max_span, span) |>
  summarise(correct = mean(correct), 
            n = n()) |>
  ggplot(aes(x = span, y = correct, col = max_span)) + 
  geom_point(aes(size = n)) + 
  geom_line()
  # geom_smooth(method = "lm", se =FALSE)

```

```{r}
mg_runs <- mg |>
  filter(correct) |>
  group_by(site, user_id, run_id, direction) |>
  summarise(max_span = max(span), 
            age = mean(age))
```


```{r}
ggplot(mg_runs, aes(x = age, y = max_span, col = direction)) + 
  geom_jitter(alpha = .5, height = .1, width = 0) + 
  geom_smooth(method = "lm")+ 
  facet_grid(direction~site, scales = "free_y")
```

## IRT

We now use the multigroup IRT models. 

```{r}
multigroup_scores_mg <- multigroup_scores |>
  filter(task_id == "memory-game", site != "us_pilot") |>
  select(site, task_id, user_id, run_id, metric_type, metric_value ) |>
  left_join(run_ages)

ggplot(multigroup_scores_mg, aes(x = age, y = metric_value, col = site)) + 
  geom_point() + 
  geom_smooth()
```

Look at coefficients. 

```{r}
mg_coefs <- best_multigroup |>
  filter(task_id == "memory-game", term == "difficulty", site != "us_pilot") |>
  separate(item, sep = "_", into = c("task", "direction", "size", "span", "number")) |>
  mutate(span = as.numeric(str_replace(span, "len", ""))) |>
  select(-task, -number) |>
  distinct()

 mg_irt_trial_type <- ggplot(mg_coefs, aes(x = span, y = -value, colour = direction)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE) + 
  facet_wrap(~site) +
  labs(
    title = "Memory Game: IRT difficulty by span and direction",
    x = "Span length",
    y = "IRT difficulty (flipped)",
    colour = "Direction"
  )

print(mg_irt_trial_type)

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "mg_irt_trial_type.png"),
  plot = mg_irt_trial_type,
  width = 10,
  height = 6,
  dpi = 300
)

# Span graph
mg_coefs <- best_multigroup |>
  filter(task_id == "memory-game", term == "difficulty", site != "us_pilot") |>
  separate(item, sep = "_", into = c("task", "direction", "size", "span", "number")) |>
  mutate(
    span = as.numeric(str_replace(span, "len", "")),
    size = recode(size, "2grid" = "2×2 grid", "3grid" = "3×3 grid")
  ) |>
  select(site, span, size, direction, value)

# one point per condition
mg_spans_irt <- mg_coefs |>
  group_by(site, span, size, direction) |>
  summarise(mean_difficulty = mean(value, na.rm = TRUE), .groups = "drop")

# Plot
mg_span_irt <- ggplot(mg_spans_irt, aes(x = span, y = -mean_difficulty, col = direction, pch = size)) +
  geom_point(size = 3) +
  geom_line(aes(lty = size)) +
  facet_wrap(~site) +
  labs(
    title = "Memory Game: IRT difficulty by span length, direction, and grid size",
    x = "Span length",
    y = "IRT difficulty (flipped)",
    colour = "Direction",
    shape = "Grid size",
    linetype = "Grid size"
  )

print(mg_span_irt)
ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "mg_span_irt.png"),
  plot = mg_span_irt,
  width = 10,
  height = 6,
  dpi = 300
)





# Mike's previous code
mg_coefs <- filter(best_multigroup, task_id == "memory-game")$coefs[[1]] |>
  separate(item, sep = "_", into = c("task", "direction", "size", "span", "number")) |>
  mutate(span = as.numeric(str_replace(span, "len",""))) |>
  select(-task, -number) |>
  distinct()

ggplot(mg_coefs, aes(x = span, y = -d, col = direction, pch = size)) + 
  geom_point() + 
  geom_line() + 
  # geom_text_repel(aes(label = direction)) + 
  xlab("Slope") + 
  ylab("Difficulty") + 
  facet_wrap(~site) 



```

Note that we have used grid3 very little and so it's getting dropped from most analysis. 


# SDS

```{r}
# remove US data
sds <- ef |>
  filter(task_id == "same-different-selection", site != "us_pilot") |>
  separate(item_uid , into = c("task", "block", "trial"), sep = "_")

```

## Sumscore

```{r}
sds_runs <- sds |>
  group_by(site, user_id, run_id, task_id) |>
  summarise(correct = mean(correct), 
            age = mean(age))

sds_runs_block <- sds |>
  filter(block != "3unique") |>
  mutate(block = fct_relevel(block, "dimensions","same")) |>
  group_by(site, user_id, run_id, task_id, block) |>
  summarise(correct = mean(correct), 
            age = mean(age))

ggplot(sds_runs_block, aes(x = age, y = correct)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se=FALSE) + 
  facet_grid(site ~ block)
```

## IRT

```{r}
multigroup_scores_sds <- multigroup_scores |>
  filter(task_id == "same-different-selection", site != "us_pilot") |>
  select(site, task_id, user_id, run_id, metric_type, metric_value ) |>
  left_join(run_ages)

ggplot(multigroup_scores_sds, aes(x = age, y = metric_value, col = site)) + 
  geom_point() + 
  geom_smooth()
```
Look at coefficients. 

```{r}

#05.05
sds_coefs <- best_multigroup |>
  filter(task_id == "same-different-selection", term == "difficulty", site != "us_pilot") |>
  separate(item, sep = "_", into = c("task", "block", "trial", "number")) |>
  filter(block != "3unique") |>
  mutate(block = fct_relevel(block, "dimensions", "same")) |>
  distinct()

# Label missing trials as "same" and classify trial type
sds_coefs_clean <- sds_coefs |>
  mutate(
    trial = if_else(trial == "", "same", trial),
    trial_type = case_when(
      str_detect(trial, "choice") ~ "choice-based",
      TRUE ~ "feature-based"
    )
  )

# Plot: All trials
ggplot(sds_coefs_clean, aes(x = block, y = -value, colour = trial)) + 
  geom_jitter(width = 0.2, height = 0.1, size = 2) +
  facet_wrap(~site) +
  coord_flip() +
  labs(
    title = "Same-Different Selection: IRT difficulty by block and trial type",
    x = "Block",
    y = "IRT difficulty (flipped)", 
    colour = "Trial type"
  )

# Subset and plot choice-based
sds_choice <- sds_coefs_clean |> 
  filter(trial_type == "choice-based") |>
  mutate(block = fct_relevel(block, "2match", "3match", "4match"))

sds_choice_plot <- ggplot(sds_choice, aes(x = block, y = -value, colour = trial)) + 
  geom_point() + 
  coord_flip() +
  facet_wrap(~site) +
  labs(
    title = "Same-different selection: Choice-based IRT difficulty by block",
    x = "Block",
    y = "IRT Difficulty (flipped)",
    colour = "Choice"
  )

print(sds_choice_plot)

ggsave(
  filename = here::here("03_explore_tasks", "Graphs", "sds_choice_plot.png"),
  plot = sds_choice_plot,
  width = 10,
  height = 6,
  dpi = 300
)

# Subset and plot feature-based !!
sds_feature <- filter(sds_coefs_clean, trial_type == "feature-based")

ggplot(sds_feature, aes(x = block, y = -value, colour = trial)) + 
  geom_point() + 
  coord_flip() +
  facet_wrap(~site) +
  labs(
    title = "Same-different selection: Feature-based IRT difficulty by block",
    x = "Block",
    y = "IRT difficulty (flipped)",
    colour = "Feature"
  )






# Mike's previous code
sds_coefs <- filter(best_multigroup, task_id == "same-different-selection")$coefs[[1]] |>
  separate(item, sep = "_", into = c("task", "block","trial")) |>
  filter(block != "3unique") |>
  mutate(block = fct_relevel(block, "dimensions","same")) |>
  distinct()

ggplot(sds_coefs, aes(x = block, y = -d, col = trial)) + 
  geom_point() + 
  geom_line(aes(group = trial)) + 
  coord_flip() +
  # geom_text_repel(aes(label = direction)) + 
  xlab("block") + 
  ylab("Difficulty") 

```

# De-bugging feature based trials

```{r}


```


# EF factor analysis

## Only LEVANTE tasks

```{r}

ef_scores <- filter(multigroup_scores, 
                    task_id %in% c("hearts-and-flowers", "memory-game", "same-different-selection"))

ef_scores_irt <- ef_scores |>
  select(site, task_id, metric_value, user_id) |>
  pivot_wider(names_from = "task_id", values_from = "metric_value", 
              id_cols = c("user_id", "site")) |>
  janitor:::clean_names() |>
  mutate(across(hearts_and_flowers:same_different_selection, ~ scale(.x)[,1]), 
         ca_site = site == "ca_pilot",
         de_site = site == "de_pilot") |>
  left_join(ages)
```

```{r}
cfa_model <-  "
ef =~ hearts_and_flowers + memory_game + same_different_selection
ef ~ age
"

fit <- cfa(cfa_model, ef_scores_irt, std.lv=TRUE, missing='fiml')
summary(fit, fit.measures=TRUE, standardize=TRUE)

graph_sem(model = fit, text_size = 3) + 
  theme(panel.background = element_rect(fill = "white"))
```

Multi-group model

```{r}
# configural invariance
fit1 <- cfa(cfa_model, ef_scores_irt, std.lv=TRUE, missing='fiml', 
           group = "site")

# weak invariance
fit2 <- cfa(cfa_model, ef_scores_irt, std.lv=TRUE, missing='fiml', 
           group = "site", group.equal = "loadings")

# strong invariance
fit3 <- cfa(cfa_model, ef_scores_irt, std.lv=TRUE, missing='fiml', 
           group = "site", group.equal = c("intercepts","loadings"))

# model comparison tests
lavTestLRT(fit1, fit2, fit3)
```

```{r}
graph_sem(model = fit1, text_size = 3) + 
  theme(panel.background = element_rect(fill = "white"))
```

## Add MEFS

```{r}
mefs_scores <- read_csv(here("02_scored_data","mefs_data","LEVANTE_20250403_1132.csv")) |>
  janitor::clean_names() |>
  select(child_id, a1_total_score, a1_standard_score) |>
  rename(user_id = child_id, 
         mefs_total = a1_total_score,
         mefs_standard = a1_standard_score
         ) 

ef_scores_irt <- ef_scores_irt |>
  left_join(mefs_scores) |>
  mutate(mefs_total = scale(mefs_total)[,1],
         mefs_standard = scale(mefs_standard)[,1])

```

```{r}
cfa_model_mefs <-  "
ef =~ hearts_and_flowers + memory_game + same_different_selection + mefs_standard
ef ~ age
"

fit <- cfa(cfa_model_mefs, ef_scores_irt, std.lv=TRUE, missing='fiml')
summary(fit, fit.measures=TRUE, standardize=TRUE)

graph_sem(model = fit, text_size = 3) + 
  theme(panel.background = element_rect(fill = "white"))
```


```{r}
# configural invariance
fit1 <- cfa(cfa_model_mefs, ef_scores_irt, std.lv=TRUE, missing='fiml', 
           group = "site")

# weak invariance
fit2 <- cfa(cfa_model_mefs, ef_scores_irt, std.lv=TRUE, missing='fiml', 
           group = "site", group.equal = "loadings")

# strong invariance
fit3 <- cfa(cfa_model_mefs, ef_scores_irt, std.lv=TRUE, missing='fiml', 
           group = "site", group.equal = c("intercepts","loadings"))

# model comparison tests
lavTestLRT(fit1, fit2, fit3)
```


```{r}
graph_sem(model = fit1, text_size = 3) + 
  theme(panel.background = element_rect(fill = "white"))
```