---
title: "LEVANTE Egma"
author: "Fionnuala O'Reilly"
toc: true
number-sections: true
highlight-style: pygments
format:
  html: 
    code-fold: true
---
Required packages

```{r}

#| echo: false

load_or_install <- function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}

load_or_install("tidyverse")
load_or_install("glue")
load_or_install("here")
load_or_install("viridis")
load_or_install("purrr")
load_or_install("readr")
load_or_install("dplyr")
load_or_install("tidyr")
load_or_install("quarto")
load_or_install("forcats")
load_or_install("ggplot2")


```

Helper functions

```{r}
#| echo: false

source(here("02_score_data","irt_helpers.R"))
source(here("03_summaries", "plotting_helper.R"))
source(here("03_explore_tasks/explore_helper.R"))

```

Get egma data

```{r}
egma <- load_task_data("egma-math")
egma
```

Exploring item group

```{r}
base::table(egma$item_group)
sort(unique(egma$item_group))

```
Sum scores

```{r}

# 94 items

colnames(egma)

egma_runs <- egma |>
  group_by(site, user_id, run_id) |>
  summarise(
    correct = mean(correct, na.rm = TRUE),  
    age = mean(age, na.rm = TRUE),          
    n_items = n_distinct(item_uid) # how many unique items were completed
  )

ggplot(egma_runs, aes(x = age, y = correct)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "re")) +
  theme_minimal(base_family = "sans") +
  ylim(0, 1) +
  facet_wrap(~site)

```

Accuracy on item groups

```{r}
egma_cat_runs <- egma |>
  group_by(site, user_id, run_id, item_group) |>
  summarise(
    correct = mean(correct, na.rm = TRUE),
    age = mean(age, na.rm = TRUE),
    n_items = n_distinct(item_uid),
    .groups = "drop"
  )

egma_counts <- egma_cat_runs |>
  group_by(site, item_group) |>
  summarise(n = n(), .groups = "drop")
```

Plot

```{r fig.height =8, fig.width = 10}

ggplot(egma_cat_runs, aes(x = age, y = correct)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  geom_text(data = egma_counts,
          aes(x = Inf, y = Inf, label = paste0("n = ", n)),
          hjust = -0.1, vjust = 1,
          size = 3, 
          inherit.aes = FALSE) +
  theme_minimal(base_family = "sans") +
  ylim(0, 1) +
  facet_grid(site ~ item_group) +
  labs(title = "EGMA Accuracy by Age, Site, and Category",
       x = "Age (years)",
       y = "Proportion Correct")

```

Cleaner graph 

```{r}

egma_type <- egma |>
  group_by(site, user_id, run_id, item_group) |>
  summarise(correct = mean(correct, na.rm = TRUE),
            age = mean(age, na.rm = TRUE))

ggplot(egma_type, aes(x = age, y = correct, col = item_group)) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal(base_family = "sans") + 
  viridis::scale_color_viridis(discrete=TRUE) +
  facet_wrap(~site)

```

Difficulty

```{r}

# Proportion correct per item per site
item_difficulty <- egma |>
  group_by(site, item_uid) |>
  summarise(p_correct = mean(correct, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = site, values_from = p_correct)

# Man accuracy across sites
egma_difficulty_raw <- item_difficulty |>
  mutate(mean_accuracy = rowMeans(across(-item_uid), na.rm = TRUE))

# Reorder items from easiest to hardest
egma_difficulty_raw <- egma_difficulty_raw |>
  mutate(item_uid = fct_reorder(item_uid, mean_accuracy))

```

Plot

```{r fig.height =20, fig.width = 10}

ggplot(egma_difficulty_raw, aes(x = item_uid, y = mean_accuracy, fill = mean_accuracy)) +
  geom_col() +
  scale_fill_viridis_c() +
  coord_flip() +
  labs(
    title = "EGMA Item Difficulty (Easiest to Hardest)",
    x = "Item (UID)",
    y = "Mean Proportion Correct",
    fill = "Accuracy"
  ) +
  ylim(0, 1) +
  theme_minimal(base_family = "sans")

# Note: The 'easiest' items are based on the proportion of correct responses, which can be misleading if only a few participants attempted them. For example, two fraction items appear among the easiest, even though fractions are generally the most difficult category—this is because the few participants who saw those items answered them correctly. IRT model accounts for this as it estimates item difficulty and discrimination while adjusting for participant ability.

```

Difficulty by item_group

```{r}
egma_groups <- egma |>
  select(item_uid, item_group) |>
  distinct()

egma_difficulty_grouped <- egma_difficulty_raw |>
  left_join(egma_groups, by = "item_uid")

# Mean accuracy by item group
egma_group_summary <- egma_difficulty_grouped |>
  group_by(item_group) |>
  summarise(
    mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n_items = n(),
    .groups = "drop"
  ) |>
  arrange(desc(mean_accuracy))

# Bar graph
ggplot(egma_group_summary, aes(x = fct_reorder(item_group, mean_accuracy), y = mean_accuracy, fill = mean_accuracy)) +
  geom_col() +
  scale_fill_viridis_c() +
  coord_flip() +
  labs(
    title = "Average Difficulty by Item Group (EGMA)",
    x = "Item Group",
    y = "Mean Proportion Correct",
    fill = "Accuracy"
  ) +
  ylim(0, 1) +
  theme_minimal(base_family = "sans")

```

Box plot - spread of accuracy across categories

```{r}

ggplot(egma_difficulty_grouped, aes(x = fct_reorder(item_group, mean_accuracy), y = mean_accuracy, fill = item_group)) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  coord_flip() +
  labs(
    title = "Distribution of Item Difficulty by Group (EGMA)",
    x = "Item Group",
    y = "Mean Accuracy"
  ) +
  theme_minimal(base_family = "sans") +
  theme(legend.position = "none")

```

Reaction times
```{r}

summary(egma$rt_numeric)

egma_trimmed <- egma |> 
  filter(rt_numeric > 200, rt_numeric < 5000)

ggplot(egma_trimmed, aes(x = rt_numeric)) +
  geom_histogram(binwidth = 100, fill = "steelblue", colour = "white") +
  facet_wrap(~site, scales = "free_y") +
  theme_minimal(base_family = "sans") +
  labs(
    title = "Histogram of Reaction Times by Site",
    x = "Reaction Time (ms)",
    y = "Count"
  )

# Check outliers
summary(egma$rt_numeric)
summary(egma_trimmed$rt_numeric)


```

Load IRT results

```{r}
best_multigroup <- readRDS(here("02_scoring_outputs", "irt_outputs", "independent_item_coefs.rds"))
multigroup_scores <- readRDS(here("02_scoring_outputs", "scores", "scores_multigroup.rds"))

```

Multigroup models

Developmental trends in egma math ability estimates by site

This plot shows how IRT-derived ability estimates (theta scores, i.e. estimated latent ability) from the Egma task vary with age across different sites.
```{r}

run_ages <- egma |>
  select(site, run_id, user_id, age) |>
  distinct()

multigroup_scores_egma <- multigroup_scores |>
  filter(task_id == "egma-math") |>
  select(site, task_id, user_id, run_id, metric_type, metric_value) |>
  left_join(run_ages)

glimpse(multigroup_scores_egma)

ggplot(multigroup_scores_egma, aes(x = age, y = metric_value, col = site)) + 
  geom_point() + 
  theme_minimal(base_family = "sans") +
  geom_smooth()

```

Get the coefficients - there doesn't seem to be any coefficients; return to this tomorrow. 

```{r}
egma_coefs <- best_multigroup |>
  filter(task_id == "egma-math", term == "difficulty", str_detect(item, "^egma_"))

head(egma_coefs)

# Calculate mean IRT difficulty across sites
irt_difficulty <- egma_coefs |>
  group_by(item) |>
  summarise(mean_d = mean(value, na.rm = TRUE), .groups = "drop")

# Join with item_group

egma_groups <- egma |>
  select(item_uid, item_group) |>
  distinct() |>
  mutate(item = glue("{item_uid}_1"))  # Adjust if item names have a different suffix

egma_difficulty_grouped <- left_join(irt_difficulty, egma_groups, by = "item")

```

Item difficulty by group

```{r}

ggplot(egma_difficulty_grouped, aes(x = fct_reorder(item_group, mean_d), y = mean_d, fill = item_group)) +
  geom_boxplot(alpha = 0.8) +
  coord_flip() +
  scale_fill_viridis_d(option = "C") +
  labs(
    title = "Distribution of EGMA Item Difficulty by Item Group",
    x = "Item Group",
    y = "Mean IRT Difficulty"
  ) +
  theme_minimal(base_family = "sans") +
  theme(legend.position = "none")
```

IRT difficulty [return to this tomorrow - april 23]

```{r}

# Calculate mean IRT difficulty (d) for each item across groups
irt_difficulty <- egma_coefs |> 
  group_by(item) |> 
  summarise(mean_d = mean(d, na.rm = TRUE)) 

# Match item to item_uid
irt_difficulty <- irt_difficulty |>
  mutate(item_uid = str_remove(item, "_1")) 

# Check
unique(irt_difficulty$item_uid)
unique(egma$item_uid)

# Join with item_group info.
irt_difficulty_grouped <- irt_difficulty |>
  left_join(egma_groups, by = "item_uid")

# Summarise mean difficulty by group
irt_group_summary <- irt_difficulty_grouped |>
  group_by(item_group) |>
  summarise(
    mean_d = mean(mean_d, na.rm = TRUE),
    n_items = n(),
    .groups = "drop"
  ) |>
  arrange(mean_d)  # Lower = easier

# Plot
ggplot(irt_group_summary, aes(x = fct_reorder(item_group, mean_d), y = mean_d, fill = mean_d)) +
  geom_col() +
  scale_fill_viridis_c() +
  coord_flip() +
  labs(
    title = "Mean IRT Difficulty by Item Group (EGMA)",
    x = "Item Group",
    y = "Mean Difficulty (d)",
    fill = "Difficulty"
  ) +
  theme_minimal(base_family = "sans")

# Return to this - line is one of the harder categories with the raw data. There's also something going on with NA. 

# END

```

```{r}

```


